### 2023-07-03 latex/presentation/main.tex
presentation init

\documentclass{beamer}
\usetheme{metropolis}

\AtBeginSection[]
{
  \begin{frame}
    \frametitle{Table of Contents}
    \tableofcontents[currentsection]
  \end{frame}
}

\title{Distributed computation of linear algebra operations}
\subtitle{Distributed systems and networks laboratory}
\author{
    Gabriele Aloisio [503264] \and
    Samuel Giacomo Raffa [matricola]
    }

\institute{Università degli studi di Messina}
\date{2023}
\logo{\includegraphics[height=1cm]{unime.png}}

\begin{document}
\maketitle

\section{Introduction}
\begin{frame}{Overview of distributed systems and networks}

\end{frame}

\begin{frame}{Importance of distributed computation in solving large-scale problems}

\end{frame}

\begin{frame}{Motivation for using Python and the Ray library}

\end{frame}

\section{Background on linear algebra operations}
\begin{frame}{Explanation of common linear algebra operations}

\end{frame}

\begin{frame}{Challenges in performing these operations on large datasets}

\end{frame}

\begin{frame}{}

\end{frame}

\section{The Ray library}
\begin{frame}{Overview of the Ray library and its capabilities}

\end{frame}

\begin{frame}{Key features and advantages of using Ray for distributed computation}

\end{frame}

\section{Implementation in Python}
\begin{frame}{List of operation}
    We will showcase the following operations:
    \begin{itemize}
        \item{Product}
        \item{Determinant}
        \item{Inverse}
        \item{Rank}
    \end{itemize}
\end{frame}

\subsection{The Matrix and RayMatrix classes}
\begin{frame}{The Matrix and RayMatrix classes}

\end{frame}

\begin{frame}{The Matrix class}

\end{frame}

\begin{frame}{The RayMatrix class}

\end{frame}

\subsection{Operations}

\end{document}

### 2023-07-03 latex/presentation/main.tex
Introduction init

@@
 \section{Introduction}
-\begin{frame}{Overview of distributed systems and networks}
-
+\begin{frame}{Matrix operations}
+    Matrix operations are fundamental in numerous scientific and computational domains, serving as the building blocks for various applications.
+However, traditional sequential computation on a single machine often becomes a bottleneck, limiting the speed and scalability of matrix calculations.
 \end{frame}
 
-\begin{frame}{Importance of distributed computation in solving large-scale problems}
-
+\begin{frame}{Matrix operations}
+    To address this problem, we are going to use an open-source distributed computing framework called \alert{Ray}.
 \end{frame}
 
-\begin{frame}{Motivation for using Python and the Ray library}
+\begin{frame}{Overview of distributed systems and networks}
+A \alert{distributed system} consists of multiple interconnected computers that collaborate and coordinate their activities to achieve a common goal.
+These systems are designed to tackle  tasks that cannot be efficiently computed by a single machine.
 
+\alert{Networks} serve as the backbone of distributed systems, enabling communication among the connected nodes.
+Network infrastructures enable coordination, data sharing, and synchronization across distributed systems, regardless of their physical locations.
+\end{frame}
+
+\begin{frame}{Issues with distributed systems}
+    Distributed systems also present unique challenges:
+    \begin{itemize}
+        \item{Data consistency}
+        \item {Fault tolerance}
+        \item {Load balancing}
+        \item {Network latency}
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Solving challenges with Ray}
+    When it comes to matrix operation computation, using Ray for distributed computation offers several advantages over performing the operations on a single machine:
+    \begin{itemize}
+    \item Faster execution
+    \item Scalability
+    \item Fault tolerance
+    \item Resource utilization
+    \end{itemize}
+\end{frame}
+
+\begin{frame}{Faster execution}
+    With Ray you can use multiple machines and CPUs to compute operations.
+This significantly reduces the computation time compared to a single machine.
+Each machine can work on a subset of the matrix data, completing the calculations in parallel.
+This distributed approach can lead to substantial speedups.
+\end{frame}
+
+\begin{frame}{Scalability}
+    As the size of the matrices grows, a single machine may struggle to handle the computational demands due to memory limitations or processing power constraints.
+Ray allows you to scale horizontally by adding more machines to the distributed setup.
+\end{frame}
+
+\begin{frame}{Fault tolerance}
+    Ray offers fault tolerance mechanisms that ensure the continuity of computation even in the presence of failures.
+If a machine participating in the distributed computation fails, Ray can automatically redistribute the workload to other available machines.
+\end{frame}
+
+\begin{frame}{Resource utilization}
+    With Ray, each machine contributes its processing power and memory capacity to the overall computation.
+This efficient utilization of resources allows you to make the most of the available hardware infrastructure, compared to a single machine that may be underutilized.
 \end{frame}

### 2023-07-04 latex/presentation/main.tex
removed section
multiplication and determinant

@@
 \usetheme{metropolis}
+\usepackage{verbatim}
 
@@
 
-\section{Background on linear algebra operations}
-\begin{frame}{Explanation of common linear algebra operations}
-
+\section{Implementation in Python}
+\begin{frame}{The Matrix and RayMatrix classes}
+    Before implementing the parallelized algorithms for Ray, we wrote the code for the serial execution in the \alert{Matrix} class:
 \end{frame}
 
-\begin{frame}{Challenges in performing these operations on large datasets}
+\begin{frame}[fragile]{The Matrix class}
+    \begin{verbatim}
+        class Matrix:
+            def __init__(self, elements):
+                [...]
+                self.elements = elements
+            def inv(self): [...]
+            def det(self): [...]
+            def product(self, b): [...]
+            def rank(self, order): [...]
+            [...]
 
+    \end{verbatim}
+
+    (Most of the functions defined in the class are auxiliary therefore were excluded for practical purposes)
 \end{frame}
 
-\begin{frame}{}
-
+\begin{frame}{The RayMatrix class}
+    The \alert{RayMatrix} class is an extension of the Matrix class that inherits its attributes and methods.
+Operations that had to be parallelized were overrided and new auxiliary methods (methods that start with \textit{task\textunderscore}) decorated with \textbf{@ray.remote} were added:
 \end{frame}
 
-\section{The Ray library}
-\begin{frame}{Overview of the Ray library and its capabilities}
+\begin{frame}[fragile]{The RayMatrix class}
+    \begin{tiny}
+        \begin{verbatim}
+        class RayMatrix(Matrix):
+            def __init__(self, elements):
+                super().__init__(elements)
 
+            @ray.remote
+            def task_rank_det(submatrix, j):
+
+            @ray.remote
+            def task_get_square_submatrix(self, start_row, start_col, row, order):
+
+            @ray.remote
+            def task_det(self, elements, i):
+
+            @ray.remote
+            def task_inv_cof(self, a, i, j):
+
+            @staticmethod
+            @ray.remote
+            def task_multiply(a, b, i, j, k):
+
+            @staticmethod
+            @ray.remote
+            def task_sum(results):
+
+    \end{verbatim}
+    \end{tiny}
 \end{frame}
 
-\begin{frame}{Key features and advantages of using Ray for distributed computation}
-
-\end{frame}
-
-\section{Implementation in Python}
+\section{Operations}
 \begin{frame}{List of operation}
@@
     \begin{itemize}
-        \item{Product}
+        \item{Multiplication}
         \item{Determinant}
@@
 
-\subsection{The Matrix and RayMatrix classes}
-\begin{frame}{The Matrix and RayMatrix classes}
+\subsection{Multiplication}
+\begin{frame}{Multiplication}
+    \alert{Matrix multiplication} is a binary operation that produces a matrix from two matrices.
+For matrix multiplication, the number of columns in the first matrix must be equal to the number of rows in the second matrix.
+The resulting matrix, known as the matrix \textbf{product}, has the number of rows of the first and the number of columns of the second matrix.
+    $$
+        A = \begin{bmatrix}
+            a_{11} & a_{12} & \dots  & a_{1n} \\
+            a_{21} & a_{22} & \dots  & a_{2n} \\
+            \vdots & \vdots & \ddots & \vdots \\
+            a_{m1} & a_{m2} & \dots  & a_{mn}
+        \end{bmatrix}, \quad
+        B = \begin{bmatrix}
+            b_{11} & b_{12} & \dots  & b_{1p} \\
+            b_{21} & b_{22} & \dots  & b_{2p} \\
+            \vdots & \vdots & \ddots & \vdots \\
+            b_{n1} & b_{n2} & \dots  & b_{np}
+        \end{bmatrix}
+    $$
+\end{frame}
+
+\begin{frame}{Multiplication}
+    The matrix multiplication of $A$ and $B$ is denoted as $C = A \cdot B$. The resulting matrix $C$ will have dimensions $m \times p$.
+
+    The entry $c_{ij}$ of the resulting matrix $C$ is computed as follows:
+
+    $$
+        c_{ij} = a_{i1} \cdot b_{1j} + a_{i2} \cdot b_{2j} + \dots + a_{in} \cdot b_{nj} = \sum_{k=1}^{n} a_{ik} \cdot b_{kj}
+    $$
+
+    The resulting matrix $C$ can be expressed as:
+
+    $$
+        C = \begin{bmatrix}
+            c_{11} & c_{12} & \dots  & c_{1p} \\
+            c_{21} & c_{22} & \dots  & c_{2p} \\
+            \vdots & \vdots & \ddots & \vdots \\
+            c_{m1} & c_{m2} & \dots  & c_{mp}
+        \end{bmatrix}
+    $$
+\end{frame}
+
+\begin{frame}[fragile]{Multiplication with Ray (1)}
+    \begin{small}
+        \begin{verbatim}
+@staticmethod
+def product(a, b):
+    a_columns = a.size()["columns"]
+    a_rows = a.size()["rows"]
+
+    b_columns = b.size()["columns"]
+
+    if a_rows != b_columns:
+        raise ValueError(
+            """
+            Number of columns of the first matrix 
+            must match the number of rows of the second matrix
+            """
+            )
+    else:
+    ...
+
+    \end{verbatim}
+    \end{small}
+\end{frame}
+
+\begin{frame}[fragile]{Multiplication with Ray (2)}
+    \begin{tiny}
+        \begin{verbatim}
+a_elements = a.get()  # Get the elements of matrix a
+b_elements = b.get()  # Get the elements of matrix b
+
+result = [[0] * b_columns for _ in range(a_rows)]  # Create a result matrix filled with zeros
+tasks = []  # Initialize an empty list to store the tasks
+
+# Iterate over rows of matrix a
+for i in range(a_rows):
+    # Iterate over columns of matrix b
+    for j in range(b_columns):
+        elements_to_multiply = []  # Initialize an empty list to store the elements to be multiplied
+
+        # Iterate over columns of matrix a and rows of matrix b
+        for k in range(a_columns):
+            # Add a task to multiply the elements at position (i, k) and (k, j)
+            elements_to_multiply.append(
+                RayMatrix.task_multiply.remote(a=a_elements, b=b_elements, i=i, j=j, k=k)
+            )
+
+        # Add the task to compute the sum of multiplied elements
+        tasks.append(RayMatrix.task_sum.remote(results=ray.get(elements_to_multiply)))
+    \end{verbatim}
+    \end{tiny}
+\end{frame}
+
+\begin{frame}[fragile]{Multiplication with Ray (3)}
+    \begin{verbatim}
+# Get the results of all the tasks
+results = ray.get(tasks)
+
+# Fill the result matrix with the computed results
+for i in range(a_rows):
+    for j in range(b_columns):
+        result[i][j] = results.pop(0)
+
+return RayMatrix(result)
+
+    \end{verbatim}
+\end{frame}
+
+\begin{frame}[fragile]{Multiplication with Ray(4)}
+    \begin{verbatim}
+        @staticmethod
+        @ray.remote
+        def task_multiply(a, b, i, j, k):
+            return a[i][k] * b[k][j]
+    \end{verbatim}
+\end{frame}
+
+\subsection{Determinant}
+\begin{frame}{Determinant}
+    The \alert{determinant} is a scalar value that is a function of the entries of a square matrix.
+It characterizes some properties of the matrix and the linear map represented by the matrix.
+In particular, the determinant is nonzero if and only if the matrix is \textbf{invertible}.
 
@@
 
-\begin{frame}{The Matrix class}
+\begin{frame}{Determinant - Laplace method}
+    To compute the determinant of matrices with order greater than 2 we are going to use the \alert{Laplace method}.
+
+    Let $A$ be a square matrix of size $n \times n$. The Laplace expansion of the determinant along the $i$th row is given by:
+
+    $$
+        \det(A) = a_{i1} C_{i1} + a_{i2} C_{i2} + \ldots + a_{in} C_{in}
+    $$
+
+    where $C_{ij}$ denotes the cofactor of the element $a_{ij}$.
+\end{frame}
+
+\begin{frame}{Determinant - Cofactor}
+    The \alert{cofactor} $C_{ij}$ is calculated as follows:
+
+    $$
+        C_{ij} = (-1)^{i+j} \cdot \det(M_{ij})
+    $$
+
+    where $\det(M_{ij})$ represents the determinant of the submatrix obtained by deleting the $i$th row and $j$th column from matrix $A$.
+\end{frame}
+
+\begin{frame}{Determinant}
+    Using the Laplace method, the determinant can be calculated recursively by expanding along any row or column until a $2 \times 2$ matrix is reached, for which the determinant can be directly computed.
+
+    The Laplace method provides an alternative approach for determining the determinant of a matrix and can be particularly useful for matrices of larger sizes.
+\end{frame}
+
+\begin{frame}[fragile]{Determinant with Ray (1)}
+    \begin{tiny}
+        \begin{verbatim}
+def det(self):
+if self.is_square():
+    size = self.size()["rows"]
+    a = self.get()
+
+    if size == 1:
+        return a[0][0]
+
+    elif size == 2:
+        return (a[0][0] * a[1][1]) - (a[0][1] * a[1][0])
+
+    else:
+        sum = 0
+
+        for i in range(1, size):
+            print(i)
+
+            futures = self.task_det.remote(self=self, elements=a, i=i)
+            print(ray.get(futures))
+
+            sum += ray.get(futures)
+
+        return sum
+
+else:
+    raise ValueError("Cannot compute determinant of a non-square matrix")
+
+            \end{verbatim}
+    \end{tiny}
+\end{frame}
+
+\begin{frame}[fragile]{Determinant with Ray (2)}
+
+    \begin{small}
+        \begin{verbatim}
+@ray.remote
+def task_det(self, elements, i):
+    size = len(elements)
+    submatrix_det_sum = 0
+    mats = self.get_square_submatrices(i)
+
+    for j in range(1, size):
+        submatrix_det = mats[j].det2()
+        submatrix_det_sum += 
+            elements[i][j] * ((-1) ** (i + j + 2)) 
+                * submatrix_det
+
+    return submatrix_det_sum
+    \end{verbatim}
+    \end{small}
 
@@
 
-\begin{frame}{The RayMatrix class}
-
-\end{frame}
-
-\subsection{Operations}
-
 \end{document}

### 2023-07-04 mat/raymatrix.py

@@ -187,7 +187,6 @@ class RayMatrix(Matrix):
         a_rows = a.size()["rows"]
 
         b_columns = b.size()["columns"]
-        b_rows = b.size()["rows"]
 
         if a_rows != b_columns:
             raise ValueError("Number of columns of the first matrix must match the number of rows of the second matrix")

### 2023-07-04 latex/presentation/main.tex
added math intros

@@
 
-\AtBeginSection[]
-{
-  \begin{frame}
-    \frametitle{Table of Contents}
-    \tableofcontents[currentsection]
-  \end{frame}
-}
+% \AtBeginSection[]
+% {
+%   \begin{frame}
+%     \frametitle{Table of Contents}
+%     \tableofcontents[currentsection]
+%   \end{frame}
+% }
 
@@
 \maketitle
+
+\begin{frame}{Table of contents}
+    \tableofcontents
+\end{frame}
 
@@
 
-\subsection{Multiplication}
+\section{Multiplication}
 \begin{frame}{Multiplication}
@@
 
-\subsection{Determinant}
+\section{Determinant}
 \begin{frame}{Determinant}
@@
 
-\begin{frame}{Determinant}
+\begin{frame}{Determinant - Laplace method}
     Using the Laplace method, the determinant can be calculated recursively by expanding along any row or column until a $2 \times 2$ matrix is reached, for which the determinant can be directly computed.
@@
     The Laplace method provides an alternative approach for determining the determinant of a matrix and can be particularly useful for matrices of larger sizes.
+
+    It's an essential operation since it's used in the other operations of this study.
 \end{frame}
@@
 
+\section{Inverse}
+\begin{frame}{Inverse of a matrix (1)}
+    The \alert{inverse of a matrix} is a fundamental concept in linear algebra.
+For a square matrix $A$, if an inverse exists, it is denoted as $A^{-1}$.
+A matrix is invertible (or non-singular) if and only if its determinant is non-zero.
+
+    The inverse of matrix $A$ satisfies the following condition:
+    $$
+        A \cdot A^{-1} = A^{-1} \cdot A = I
+    $$
+    where $I$ represents the identity matrix.
+
+\end{frame}
+
+\begin{frame}{Inverse of a matrix (2)}
+    To find the inverse of a matrix, one common method is to use the formula:
+    $$
+        A^{-1} = \frac{1}{\det(A)} \cdot \text{adj}(A)
+    $$
+    where $\det(A)$ denotes the determinant of matrix $A$ and $\text{adj}(A)$ represents the adjugate (or adjoint) of matrix $A$.
+
+    The adjugate of matrix $A$ is obtained by taking the transpose of the matrix of cofactors of $A$. The cofactor $C_{ij}$ is calculated as:
+    $$
+        C_{ij} = (-1)^{i+j} \cdot \det(M_{ij})
+    $$
+    where $\det(M_{ij})$ represents the determinant of the submatrix obtained by deleting the $i$th row and $j$th column from matrix $A$.
+\end{frame}
+
+\begin{frame}[fragile]{Inverse with Ray}
+    \begin{small}
+        \begin{verbatim}
+            codice inversa
+        \end{verbatim}
+    \end{small}
+\end{frame}
+
+\section{Rank}
+\begin{frame}{Rank (1)}
+    The \alert{rank} of a matrix $A$, denoted as $\text{rank}(A)$, is defined as the maximum number of linearly independent rows or columns in the matrix.
+
+    The rank of a matrix can be determined using the minor criterion, also known as the \alert{criterion of minors}.
+According to this criterion, the rank of a matrix is equal to the largest order of a non-zero determinant of any square submatrix within the given matrix.
+\end{frame}
+
+\begin{frame}{Criterion of minors}
+
+Let $A$ be a matrix of size $m \times n$, and let $j$ be the order of the largest non-zero determinant among all square submatrices of $A$.
+Then the rank of $A$, denoted as $\text{rank}(A)$, is equal to $j$.
+
+To apply the minor criterion, we calculate the determinants of all possible square submatrices of $A$, ranging from $1 \times 1$ to $\min(m,n) \times \min(m,n)$.
+The order of the largest non-zero determinant among these submatrices gives us the rank of $A$.
+
+If there is a non-zero determinant of order $j$, but all determinants of order $k+1$ or higher are zero, then the rank of $A$ is $j$.
+\end{frame}
+
+\begin{frame}[fragile]{Rank with Ray}
+    \begin{small}
+        \begin{verbatim}
+            codice rango
+        \end{verbatim}
+    \end{small}
+\end{frame}
+
 \end{document}

### 2023-07-05 main.py
fix prodotto scalare, inversa e altro

@@ -1,4 +1,6 @@
 from mat.raymatrix import RayMatrix
+from mat.matrix import Matrix
+
 import ray
 import random
 
@@ -37,8 +39,18 @@ def main():
         [5, 8, 4, 1],
         [6, 3, 0, 2]
     ])
+    F = Matrix([
+        [4, 10, 2, 6],
+        [8, 9, 1, 3],
+        [5, 8, 4, 1],
+        [6, 3, 0, 2]
+    ])
+
+    print(E.inv())
+    print(F.inv())
+
     #print(RayMatrix.product(B,C))
-    print(E.det())
+    #print(E.det())
 
     ray.shutdown()
 

### 2023-07-05 mat/matrix.py

@@ -66,46 +66,46 @@ class Matrix:
 
         return Matrix(elems)
 
-    # def inv(self):
-    #     if not self.is_square():
-    #         raise Exception("Matrix must be square")
+    def inv(self):
+        if not self.is_square():
+            raise Exception("Matrix must be square")
 
-    #     elif Matrix.det(self) == 0:
-    #         raise Exception("Matrix is not invertible")
+        elif Matrix.det(self) == 0:
+            raise Exception("Matrix is not invertible")
 
-    #     else:
-    #         size = self.size()["rows"]
-    #         a = self.get()
-    #         cof_arr = []
+        else:
+            size = self.size()["rows"]
+            a = self.get()
+            cof_arr = [[0] * size for _ in range(size)]
 
-    #         for i in range(size):
-    #             for j in range(size):
-    #                 cof_arr[i][j] = a[i][j] * ((-1) ** (i + j + 2)) * self.minor(i, j).det()
+            for i in range(size):
+                for j in range(size):
+                    cof_arr[i][j] = a[i][j] * ((-1) ** (i + j + 2)) * self.minor(i, j).det()
 
-    #         cof_matrix = Matrix(cof_arr)
+            cof_matrix = Matrix(cof_arr)
 
-    #         return cof_matrix.scalar_product(1 / self.det())
+            return cof_matrix.scalar_product(1 / self.det())
 
-    # def det(self):
-    #     if self.is_square():
-    #         size = self.size()["rows"]
-    #         a = self.get()
+    def det(self):
+        if self.is_square():
+            size = self.size()["rows"]
+            a = self.get()
 
-    #         if size == 1:
-    #             return a[0][0]
+            if size == 1:
+                return a[0][0]
 
-    #         elif size == 2:
-    #             return (a[0][0] * a[1][1]) - (a[0][1] * a[1][0])
+            elif size == 2:
+                return (a[0][0] * a[1][1]) - (a[0][1] * a[1][0])
 
-    #         else:
+            else:
 
-    #             for i in range(size):
-    #                 sum = 0
+                for i in range(size):
+                    sum = 0
 
-    #                 for j in range(size):
-    #                     sum += a[i][j] * ((-1) ** (i + j + 2)) * self.minor(i, j).det()
+                    for j in range(size):
+                        sum += a[i][j] * ((-1) ** (i + j + 2)) * self.minor(i, j).det()
 
-    #             return sum
+                return sum
 
     def product(self, b):
         a_columns = self.size()["columns"]

### 2023-07-05 mat/raymatrix.py

@@ -1,6 +1,7 @@
 from mat.matrix import Matrix
 import ray
 
+
 class RayMatrix(Matrix):
     def __init__(self, elements):
         super().__init__(elements)
@@ -24,6 +25,14 @@ class RayMatrix(Matrix):
             submatrix_det_sum += elements[i][j] * ((-1) ** (i + j + 2)) * submatrix_det
 
         return submatrix_det_sum
+
+    @staticmethod
+    @ray.remote
+    def task_scalar_product(row, scalar):
+        for j in range(len(row)):
+            row[j] *= scalar
+
+        return row
 
     def get_square_submatrices(self, order):
         rows = self.size()["rows"]
@@ -168,7 +177,7 @@ class RayMatrix(Matrix):
             for i, j, value in results:
                 cof_arr[i][j] = value
 
-            cof_matrix = Matrix(cof_arr)
+            cof_matrix = RayMatrix(cof_arr)
             return cof_matrix.scalar_product(1 / self.det())
 
     @staticmethod
@@ -216,3 +225,16 @@ class RayMatrix(Matrix):
                     result[i][j] = results.pop(0)
 
             return RayMatrix(result)
+
+    def scalar_product(self, scalar):
+        elems = self.get()
+        res = []
+
+        futures = []
+
+        for r in elems:
+            futures.append(RayMatrix.task_scalar_product.remote(r, scalar))
+
+        res = ray.get(futures)
+
+        return Matrix(res)

### 2023-07-05 main.py
fix determinante parallelo

@@ -46,12 +46,14 @@ def main():
         [6, 3, 0, 2]
     ])
 
-    print(E.inv())
-    print(F.inv())
+    #print(E.inv())
+    #print(F.inv())
 
     #print(RayMatrix.product(B,C))
     #print(E.det())
 
+    print(F.det())
+    print(E.det())
     ray.shutdown()
 
 if __name__ == "__main__":

### 2023-07-05 mat/raymatrix.py

@@ -16,15 +16,18 @@ class RayMatrix(Matrix):
         return self.elements[start_row + row][start_col:start_col + order]
 
     @ray.remote
-    def task_det(self, elements, i):
-        size = len(elements)
-        submatrix_det_sum = 0
+    def task_det(self, col_index):
+        elems = self.get()
+        submat = self.minor(0, col_index)
 
-        for j in range(size):
-            submatrix_det = self.minor(i=i, j=j).det()
-            submatrix_det_sum += elements[i][j] * ((-1) ** (i + j + 2)) * submatrix_det
+        submat_det_sum = 0
 
-        return submatrix_det_sum
+        for i in range(len(elems)):
+            submatrix_det = submat.det()
+
+            submat_det_sum += elems[i][col_index] * ((-1) ** (i + col_index + 2)) * submatrix_det
+
+        return submat_det_sum
 
     @staticmethod
     @ray.remote
@@ -58,74 +61,26 @@ class RayMatrix(Matrix):
 
     def det(self):
         if self.is_square():
-            size = self.size()["rows"]
+            cols = self.size()["columns"]
             a = self.get()
 
-            if size == 1:
+            if cols == 1:
                 return a[0][0]
 
-            elif size == 2:
+            elif cols == 2:
                 return (a[0][0] * a[1][1]) - (a[0][1] * a[1][0])
 
             else:
-                '''
-                NON FUNZIONA
-                sum = 0
+                futures = []
 
-                                for i in range(size):
-                                    print(i)
+                for j in range(cols):
+                    futures.append(self.task_det.remote(self=self, col_index=j))
 
-                                    futures = self.task_det.remote(self=self, elements=a, i=0)
-                                    print(ray.get(futures))
-
-                                    sum += ray.get(futures)
-
-                                return sum'''
-
-                futures = self.task_det.remote(self=self, elements=a, i=0)
-                return ray.get(futures)
-        
-        else:
-            raise ValueError("Cannot compute determinant of a non-square matrix")
-
-    def det2(self):
-        if self.is_square():
-            size = self.size()["rows"]
-            a = self.get()
-
-            if size == 1:
-                return a[0][0]
-
-            elif size == 2:
-                return (a[0][0] * a[1][1]) - (a[0][1] * a[1][0])
-
-            else:
-                sum = 0
-
-                for i in range(1, size):
-                    print(i)
-
[…]

### 2023-07-05 main.py
inizializzato parser comandi per lettura input da shell

@@ -1,61 +1,14 @@
-from mat.raymatrix import RayMatrix
-from mat.matrix import Matrix
-
-import ray
-import random
+from startup.cmd import get_args, init_parser
 
 
-# def main():
-#     counters = [rm.ParallelMatrix.remote([
-#         [4, 5, 7],
-#         [5, 3, 2],
-#         [9, 4, 6],
-#         [1, 3, 8]
-#     ]) for i in range(4)]
-
-#     futures = [c.range.remote() for c in counters]
-#     print(ray.get(futures))
+def boot():
+    init_parser()
+    args = get_args()
 
 
-def main():
-    ray.init()
-    A = RayMatrix([
-        [1, 2, 3],
-        [4, 5, 7],
-        [5, 3, 2],
-        [9, 5, 6],
-        [1, 3, 8]])
-    
-    B = RayMatrix([[random.randint(0, 99) for _ in range(5)] for _ in range(5)])
-    C = RayMatrix([[random.randint(0, 99) for _ in range(40)] for _ in range(40)])
+def run():
+    pass
 
-    D = RayMatrix([[1,2,3],
-                  [4,5,6],
-                  [7,8,9]])
-
-    E = RayMatrix([
-        [4, 10, 2, 6],
-        [8, 9, 1, 3],
-        [5, 8, 4, 1],
-        [6, 3, 0, 2]
-    ])
-    F = Matrix([
-        [4, 10, 2, 6],
-        [8, 9, 1, 3],
-        [5, 8, 4, 1],
-        [6, 3, 0, 2]
-    ])
-
-    #print(E.inv())
-    #print(F.inv())
-
-    #print(RayMatrix.product(B,C))
-    #print(E.det())
-
-    print(F.det())
-    print(E.det())
-    ray.shutdown()
 
 if __name__ == "__main__":
-    main()
-
+    boot()

### 2023-07-05 startup/cmd.py

@@ -1,0 +1,15 @@
+from argparse import ArgumentParser
+from startup.cmdtype import unsigned
+
+parser = ArgumentParser(
+    prog="Progetto di Laboratorio di reti e sistemi distribuiti",
+    description="Progetto creato per Laboratorio di reti e sistemi distribuiti")
+
+
+def init_parser():
+    parser.add_argument('-i', '--ignore', nargs="*", choices=["p", "d", "i"], help="ignora un'operazione matriciale[p: prodotto, d: determinante, i: inversa]")
+    parser.add_argument('-l', '--load', nargs=1, type=unsigned, required=True)
+
+
+def get_args():
+    return parser.parse_args()

### 2023-07-05 startup/cmdtype.py

@@ -1,0 +1,8 @@
+from argparse import ArgumentTypeError
+
+
+def unsigned(x):
+    x = int(x)
+    if x < 0:
+        raise ArgumentTypeError("Minimum bandwidth is 12")
+    return x

### 2023-07-05 startup/cmdtype.py
fix

@@ -4,5 +4,5 @@ from argparse import ArgumentTypeError
 def unsigned(x):
     x = int(x)
     if x < 0:
-        raise ArgumentTypeError("Minimum bandwidth is 12")
+        raise ArgumentTypeError("Il valore inserito deve essere > 0")
     return x

### 2023-07-05 main.py
fix

@@ -1,4 +1,5 @@
 from startup.cmd import get_args, init_parser
+import ray
 
 
 def boot():
@@ -7,6 +8,9 @@ def boot():
 
 
 def run():
+    ray.init()
+    ray.shutdown()
+
     pass
 
 

### 2023-07-05 test.py
esecuzione test

@@ -1,0 +1,62 @@
+from mat.raymatrix import RayMatrix
+from mat.matrix import Matrix
+
+import ray
+import random
+
+
+# def main():
+#     counters = [rm.ParallelMatrix.remote([
+#         [4, 5, 7],
+#         [5, 3, 2],
+#         [9, 4, 6],
+#         [1, 3, 8]
+#     ]) for i in range(4)]
+
+#     futures = [c.range.remote() for c in counters]
+#     print(ray.get(futures))
+
+
+def main():
+    ray.init()
+    A = RayMatrix([
+        [1, 2, 3],
+        [4, 5, 7],
+        [5, 3, 2],
+        [9, 5, 6],
+        [1, 3, 8]])
+
+    B = RayMatrix([[random.randint(0, 99) for _ in range(5)] for _ in range(5)])
+    C = RayMatrix([[random.randint(0, 99) for _ in range(40)] for _ in range(40)])
+
+    D = RayMatrix([[1, 2, 3],
+                   [4, 5, 6],
+                   [7, 8, 9]])
+
+    E = RayMatrix([
+        [4, 10, 2, 6],
+        [8, 9, 1, 3],
+        [5, 8, 4, 1],
+        [6, 3, 0, 2]
+    ])
+    F = Matrix([
+        [4, 10, 2, 6],
+        [8, 9, 1, 3],
+        [5, 8, 4, 1],
+        [6, 3, 0, 2]
+    ])
+
+    # print(E.inv())
+    # print(F.inv())
+
+    # print(RayMatrix.product(B,C))
+    # print(E.det())
+
+    print(F.det())
+    print(E.det())
+    ray.shutdown()
+
+
+if __name__ == "__main__":
+    main()
+

### 2023-07-05 src/main.py
riordinata struttura del progetto

@@ -1,0 +1,18 @@
+from startup.cmd import get_args, init_parser
+import ray
+
+
+def boot():
+    init_parser()
+    args = get_args()
+
+
+def run():
+    ray.init()
+    ray.shutdown()
+
+    pass
+
+
+if __name__ == "__main__":
+    boot()

### 2023-07-05 src/mat/base_ops.py

@@ -1,0 +1,6 @@
+def get_submatrices_rank(matrix, order):
+    for m in matrix.get_square_submatrices(order):
+        if m.det() != 0:
+            return order
+
+    return -1

### 2023-07-05 src/mat/matrix.py

@@ -1,0 +1,162 @@
+from mat import base_ops as bm
+
+
+class Matrix:
+    def __init__(self, elements):
+        for i in elements:
+            if len(i) != len(elements[0]):
+                raise ValueError("Invalid row size")
+
+        self.elements = elements
+
+    def __str__(self):
+        tmp = "\n"
+
+        max_lengths = []
+        for i in range(len(self.elements[0])):
+            column_lengths = [len(str(row[i])) for row in self.elements]
+            max_lengths.append(max(column_lengths))
+
+        for row in self.elements:
+            tmp += "| "
+            for i in range(len(self.elements[0])):
+                tmp += f"{str(row[i]):^{max_lengths[i]}} "
+            tmp += "|\n"
+
+        return tmp
+
+    def size(self):
+        rows = len(self.elements)
+        columns = len(self.elements[0])
+
+        return {"rows": rows, "columns": columns}
+
+    def get(self):
+        return self.elements
+
+    def transpose(self):
+        a = self.get()
+        rows = self.size()["rows"]
+        columns = self.size()["columns"]
+
+        transpose_matrix = [[0] * rows for _ in range(columns)]
+
+        for i in range(rows):
+            for j in range(columns):
+                transpose_matrix[j][i] = a[i][j]
+
+        return Matrix(transpose_matrix)
+
+    def is_square(self):
+        return self.size()["rows"] == self.size()["columns"]
+
+    def minor(self, i, j):
+        minor_elements = [row[:j] + row[j + 1:] for row_idx, row in enumerate(self.elements) if row_idx != i]
+        return Matrix(minor_elements)
+
+    def scalar_product(self, scalar):
+        elems = self.get()
+
+        rows = self.size()['rows']
+        cols = self.size()['columns']
+
+        for i in range(rows):
+            for j in range(cols):
+                elems[i][j] *= scalar
+
+        return Matrix(elems)
+
+    def inv(self):
+        if not self.is_square():
+            raise Exception("Matrix must be square")
+
+        elif Matrix.det(self) == 0:
+            raise Exception("Matrix is not invertible")
+
+        else:
+            size = self.size()["rows"]
+            a = self.get()
+            cof_arr = [[0] * size for _ in range(size)]
+
+            for i in range(size):
+                for j in range(size):
+                    cof_arr[i][j] = a[i][j] * ((-1) ** (i + j + 2)) * self.minor(i, j).det()
+
[…]

### 2023-07-05 src/mat/raymatrix.py

@@ -1,0 +1,195 @@
+from mat.matrix import Matrix
+import ray
+
+
+class RayMatrix(Matrix):
+    def __init__(self, elements):
+        super().__init__(elements)
+
+    @ray.remote
+    def task_rank_det(submatrix, j):
+        if submatrix.det() != 0:
+            return j
+    
+    @ray.remote
+    def task_get_square_submatrix(self, start_row, start_col, row, order):
+        return self.elements[start_row + row][start_col:start_col + order]
+
+    @ray.remote
+    def task_det(self, col_index):
+        elems = self.get()
+        submat = self.minor(0, col_index)
+
+        submat_det_sum = 0
+
+        for i in range(len(elems)):
+            submatrix_det = submat.det()
+
+            submat_det_sum += elems[i][col_index] * ((-1) ** (i + col_index + 2)) * submatrix_det
+
+        return submat_det_sum
+
+    @staticmethod
+    @ray.remote
+    def task_scalar_product(row, scalar):
+        for j in range(len(row)):
+            row[j] *= scalar
+
+        return row
+
+    def get_square_submatrices(self, order):
+        rows = self.size()["rows"]
+        cols = self.size()["columns"]
+        submatrices = []
+
+        for start_row in range(rows - order + 1):
+            for start_col in range(cols - order + 1):
+                submatrix = []
+                
+                for row in range(order):
+                    futures = self.task_get_square_submatrix.remote(self=self, start_row=start_row, start_col=start_col, row=row, order=order)
+                    
+                    submatrix.append(ray.get(futures))
+                
+                submatrices.append(RayMatrix(submatrix))
+
+        return submatrices
+
+    def minor(self, i, j):
+        minor_elements = [row[:j] + row[j + 1:] for row_idx, row in enumerate(self.elements) if row_idx != i]
+        return RayMatrix(minor_elements)
+
+    def det(self):
+        if self.is_square():
+            cols = self.size()["columns"]
+            a = self.get()
+
+            if cols == 1:
+                return a[0][0]
+
+            elif cols == 2:
+                return (a[0][0] * a[1][1]) - (a[0][1] * a[1][0])
+
+            else:
+                futures = []
+
+                for j in range(cols):
+                    futures.append(self.task_det.remote(self=self, col_index=j))
+
+                return sum(ray.get(futures))
+
+        else:
+            raise ValueError("Cannot compute determinant of a non-square matrix")
+
+    def rank(self):
[…]

### 2023-07-05 src/raytest.py

@@ -1,0 +1,23 @@
+from time import sleep
+import ray
+
+ray.init()
+
+sleep(10)
+
+@ray.remote
+def cube(n):
+    return n * n * n
+
+@ray.remote
+def factorial(n):
+    tmp = 1
+    
+    for i in range(1, n):
+        tmp *= i
+
+    return tmp
+
+futures = [factorial.remote(i) for i in range(20, 50)]
+
+print(ray.get(futures))

### 2023-07-05 src/startup/cmdtype.py

@@ -1,0 +1,8 @@
+from argparse import ArgumentTypeError
+
+
+def unsigned(x):
+    x = int(x)
+    if x < 0:
+        raise ArgumentTypeError("Il valore inserito deve essere > 0")
+    return x

### 2023-07-05 main.py
fix

@@ -1,18 +1,0 @@
-from startup.cmd import get_args, init_parser
-import ray
-
-
-def boot():
-    init_parser()
-    args = get_args()
-
-
-def run():
-    ray.init()
-    ray.shutdown()
-
-    pass
-
-
-if __name__ == "__main__":
-    boot()

### 2023-07-05 mat/base_ops.py

@@ -1,6 +1,0 @@
-def get_submatrices_rank(matrix, order):
-    for m in matrix.get_square_submatrices(order):
-        if m.det() != 0:
-            return order
-
-    return -1

### 2023-07-05 mat/matrix.py

@@ -1,162 +1,0 @@
-from mat import base_ops as bm
-
-
-class Matrix:
-    def __init__(self, elements):
-        for i in elements:
-            if len(i) != len(elements[0]):
-                raise ValueError("Invalid row size")
-
-        self.elements = elements
-
-    def __str__(self):
-        tmp = "\n"
-
-        max_lengths = []
-        for i in range(len(self.elements[0])):
-            column_lengths = [len(str(row[i])) for row in self.elements]
-            max_lengths.append(max(column_lengths))
-
-        for row in self.elements:
-            tmp += "| "
-            for i in range(len(self.elements[0])):
-                tmp += f"{str(row[i]):^{max_lengths[i]}} "
-            tmp += "|\n"
-
-        return tmp
-
-    def size(self):
-        rows = len(self.elements)
-        columns = len(self.elements[0])
-
-        return {"rows": rows, "columns": columns}
-
-    def get(self):
-        return self.elements
-
-    def transpose(self):
-        a = self.get()
-        rows = self.size()["rows"]
-        columns = self.size()["columns"]
-
-        transpose_matrix = [[0] * rows for _ in range(columns)]
-
-        for i in range(rows):
-            for j in range(columns):
-                transpose_matrix[j][i] = a[i][j]
-
-        return Matrix(transpose_matrix)
-
-    def is_square(self):
-        return self.size()["rows"] == self.size()["columns"]
-
-    def minor(self, i, j):
-        minor_elements = [row[:j] + row[j + 1:] for row_idx, row in enumerate(self.elements) if row_idx != i]
-        return Matrix(minor_elements)
-
-    def scalar_product(self, scalar):
-        elems = self.get()
-
-        rows = self.size()['rows']
-        cols = self.size()['columns']
-
-        for i in range(rows):
-            for j in range(cols):
-                elems[i][j] *= scalar
-
-        return Matrix(elems)
-
-    def inv(self):
-        if not self.is_square():
-            raise Exception("Matrix must be square")
-
-        elif Matrix.det(self) == 0:
-            raise Exception("Matrix is not invertible")
-
-        else:
-            size = self.size()["rows"]
-            a = self.get()
-            cof_arr = [[0] * size for _ in range(size)]
-
-            for i in range(size):
-                for j in range(size):
-                    cof_arr[i][j] = a[i][j] * ((-1) ** (i + j + 2)) * self.minor(i, j).det()
-
[…]

### 2023-07-05 mat/raymatrix.py

@@ -1,195 +1,0 @@
-from mat.matrix import Matrix
-import ray
-
-
-class RayMatrix(Matrix):
-    def __init__(self, elements):
-        super().__init__(elements)
-
-    @ray.remote
-    def task_rank_det(submatrix, j):
-        if submatrix.det() != 0:
-            return j
-    
-    @ray.remote
-    def task_get_square_submatrix(self, start_row, start_col, row, order):
-        return self.elements[start_row + row][start_col:start_col + order]
-
-    @ray.remote
-    def task_det(self, col_index):
-        elems = self.get()
-        submat = self.minor(0, col_index)
-
-        submat_det_sum = 0
-
-        for i in range(len(elems)):
-            submatrix_det = submat.det()
-
-            submat_det_sum += elems[i][col_index] * ((-1) ** (i + col_index + 2)) * submatrix_det
-
-        return submat_det_sum
-
-    @staticmethod
-    @ray.remote
-    def task_scalar_product(row, scalar):
-        for j in range(len(row)):
-            row[j] *= scalar
-
-        return row
-
-    def get_square_submatrices(self, order):
-        rows = self.size()["rows"]
-        cols = self.size()["columns"]
-        submatrices = []
-
-        for start_row in range(rows - order + 1):
-            for start_col in range(cols - order + 1):
-                submatrix = []
-                
-                for row in range(order):
-                    futures = self.task_get_square_submatrix.remote(self=self, start_row=start_row, start_col=start_col, row=row, order=order)
-                    
-                    submatrix.append(ray.get(futures))
-                
-                submatrices.append(RayMatrix(submatrix))
-
-        return submatrices
-
-    def minor(self, i, j):
-        minor_elements = [row[:j] + row[j + 1:] for row_idx, row in enumerate(self.elements) if row_idx != i]
-        return RayMatrix(minor_elements)
-
-    def det(self):
-        if self.is_square():
-            cols = self.size()["columns"]
-            a = self.get()
-
-            if cols == 1:
-                return a[0][0]
-
-            elif cols == 2:
-                return (a[0][0] * a[1][1]) - (a[0][1] * a[1][0])
-
-            else:
-                futures = []
-
-                for j in range(cols):
-                    futures.append(self.task_det.remote(self=self, col_index=j))
-
-                return sum(ray.get(futures))
-
-        else:
-            raise ValueError("Cannot compute determinant of a non-square matrix")
-
-    def rank(self):
[…]

### 2023-07-05 startup/cmd.py

@@ -1,15 +1,0 @@
-from argparse import ArgumentParser
-from startup.cmdtype import unsigned
-
-parser = ArgumentParser(
-    prog="Progetto di Laboratorio di reti e sistemi distribuiti",
-    description="Progetto creato per Laboratorio di reti e sistemi distribuiti")
-
-
-def init_parser():
-    parser.add_argument('-i', '--ignore', nargs="*", choices=["p", "d", "i"], help="ignora un'operazione matriciale[p: prodotto, d: determinante, i: inversa]")
-    parser.add_argument('-l', '--load', nargs=1, type=unsigned, required=True)
-
-
-def get_args():
-    return parser.parse_args()

### 2023-07-05 startup/cmdtype.py

@@ -1,8 +1,0 @@
-from argparse import ArgumentTypeError
-
-
-def unsigned(x):
-    x = int(x)
-    if x < 0:
-        raise ArgumentTypeError("Il valore inserito deve essere > 0")
-    return x

### 2023-07-05 test.py

@@ -1,62 +1,0 @@
-from mat.raymatrix import RayMatrix
-from mat.matrix import Matrix
-
-import ray
-import random
-
-
-# def main():
-#     counters = [rm.ParallelMatrix.remote([
-#         [4, 5, 7],
-#         [5, 3, 2],
-#         [9, 4, 6],
-#         [1, 3, 8]
-#     ]) for i in range(4)]
-
-#     futures = [c.range.remote() for c in counters]
-#     print(ray.get(futures))
-
-
-def main():
-    ray.init()
-    A = RayMatrix([
-        [1, 2, 3],
-        [4, 5, 7],
-        [5, 3, 2],
-        [9, 5, 6],
-        [1, 3, 8]])
-
-    B = RayMatrix([[random.randint(0, 99) for _ in range(5)] for _ in range(5)])
-    C = RayMatrix([[random.randint(0, 99) for _ in range(40)] for _ in range(40)])
-
-    D = RayMatrix([[1, 2, 3],
-                   [4, 5, 6],
-                   [7, 8, 9]])
-
-    E = RayMatrix([
-        [4, 10, 2, 6],
-        [8, 9, 1, 3],
-        [5, 8, 4, 1],
-        [6, 3, 0, 2]
-    ])
-    F = Matrix([
-        [4, 10, 2, 6],
-        [8, 9, 1, 3],
-        [5, 8, 4, 1],
-        [6, 3, 0, 2]
-    ])
-
-    # print(E.inv())
-    # print(F.inv())
-
-    # print(RayMatrix.product(B,C))
-    # print(E.det())
-
-    print(F.det())
-    print(E.det())
-    ray.shutdown()
-
-
-if __name__ == "__main__":
-    main()
-

### 2023-07-05 src/main.py
configurazione calcolo

@@ -1,17 +1,26 @@
 from startup.cmd import get_args, init_parser
+import utils.config as cfg
 import ray
+import process.processor as proc
+import utils.dict_utils as du
 
 
 def boot():
     init_parser()
     args = get_args()
 
+    cfg.MATRIX_SIZE = args.load
+    
+    proc.operations = du.filter(proc.operations, args.ignore)
+
+    run()
+
 
 def run():
     ray.init()
-    ray.shutdown()
 
-    pass
+    #processor.run_test()
+    ray.shutdown()
 
 
 if __name__ == "__main__":

### 2023-07-05 src/process/generator.py

@@ -1,0 +1,2 @@
+def generate_matrix(rows: int, cols: int):
+    pass

### 2023-07-05 src/process/processor.py

@@ -1,0 +1,15 @@
+def product(a, b):
+
+    pass
+
+def determinant(a):
+    pass
+
+def inverse(a):
+    pass
+
+operations = {
+    "p": product,
+    "d": determinant,
+    "i": inverse
+}

### 2023-07-05 src/utils/config.py

@@ -1,0 +1,1 @@
+MATRIX_SIZE = 0

### 2023-07-05 src/utils/dict_utils.py

@@ -1,0 +1,5 @@
+def filter(a, keys):
+    for e in keys:
+        del a[e]
+
+    return a

### 2023-07-05 src/main.py
fix, si deve sistemare la generazione della matrice

@@ -9,7 +9,7 @@ def boot():
     init_parser()
     args = get_args()
 
-    cfg.MATRIX_SIZE = args.load
+    cfg.MATRIX_SIZE = int(args.load[0])
     
     proc.operations = du.filter(proc.operations, args.ignore)
 
@@ -19,7 +19,8 @@ def boot():
 def run():
     ray.init()
 
-    #processor.run_test()
+    proc.exec_test(cfg.MATRIX_SIZE)
+
     ray.shutdown()
 
 

### 2023-07-05 src/process/generator.py

@@ -1,2 +1,11 @@
-def generate_matrix(rows: int, cols: int):
-    pass
+import random
+
+def generate_matrix(rows, cols):
+    mat = []
+    print(rows, cols)
+    
+    for i in range(rows):
+        for j in range(cols):
+            mat[i][j] = random.randint(0, 99)
+
+    return mat

### 2023-07-05 src/process/processor.py

@@ -1,12 +1,38 @@
-def product(a, b):
+from process.generator import generate_matrix
+import utils.config as cfg
+from mat.matrix import Matrix
+from mat.raymatrix import RayMatrix
+from time import time
 
-    pass
+def exec_test(size):
+    a = generate_matrix(size, size)
+    b = generate_matrix(size, size)
+    
+    a_base = Matrix(a)
+    a_ray = RayMatrix(a)
+
+    b_base = Matrix(b)
+    b_ray = RayMatrix(b)
+
+    for op in operations():
+        exec_time_single = op(a_base, b_base)
+        exec_time_parallel = op(a_ray, b_ray)
+
+        print('time single: ', exec_time_single)
+        print('time parallel: ', exec_time_parallel)
+    
+def product(a, b):
+    start = time()
+    
+    return time() - start()
 
 def determinant(a):
     pass
 
 def inverse(a):
+
     pass
+
 
 operations = {
     "p": product,

### 2023-07-05 src/startup/cmdtype.py

@@ -1,7 +1,7 @@
 from argparse import ArgumentTypeError
 
 
-def unsigned(x):
+def unsigned(x): 
     x = int(x)
     if x < 0:
         raise ArgumentTypeError("Il valore inserito deve essere > 0")

### 2023-07-06 scripts/plot.py
aggiunto salvataggio dei dati su file csv, completato main

@@ -1,0 +1,38 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import csv
+import math
+
+from src.data.storage_manager import load_data
+
+mongo = []
+neo = []
+queryn = 4
+
+for i in range(25, 101, 25):
+    single_data = load_data('single')
+    parallel_data = load_data('parallel')
+
+
+labels = ['25%', '50%', '75%', '100%']
+x = np.arange(len(labels))  # the label locations
+width = 0.35  # the width of the bars
+
+fig, ax = plt.subplots()
+
+rects1 = ax.bar(x - width / 2, neo, width, label='Neo4j')
+rects2 = ax.bar(x + width / 2, mongo, width, label='MongoDB')
+
+# Add some text for labels, title and custom x-axis tick labels, etc.
+
+ax.set_ylabel('Tempo (Millisecondi)')
+ax.set_title('3 WHERE 2 JOIN')
+ax.set_xticks(x, labels)
+ax.legend()
+
+ax.bar_label(rects1, padding=3)
+ax.bar_label(rects2, padding=3)
+
+fig.tight_layout()
+
+plt.show()

### 2023-07-06 src/data/storage_manager.py

@@ -1,0 +1,21 @@
+from utils.data_utils import data_path
+from utils.dict_utils import csv_to_dict
+import csv
+
+
+def store_data(filename, data_model):
+    with open(data_path(filename + '.csv'), "w", newline="") as fq:
+        writer = csv.writer(fq)
+        writer.writerow(['Run type:', filename.capitalize()])
+
+        for op, time in data_model.items():
+            writer.writerow([op, time])
+
+        fq.close()
+
+
+def load_data(filename):
+    with open(data_path(filename + '.csv'), 'w', newline="") as fq:
+        data = csv_to_dict(csv_to_dict(list(csv.DictReader(fq))))
+
+        return data

### 2023-07-06 src/main.py

@@ -3,15 +3,15 @@ import utils.config as cfg
 import ray
 import process.processor as proc
 import utils.dict_utils as du
-
-
+from data.storage_manager import store_data
 def boot():
     init_parser()
     args = get_args()
 
     cfg.MATRIX_SIZE = int(args.load[0])
-    
-    proc.operations = du.filter(proc.operations, args.ignore)
+
+    if args.ignore is not None:
+        proc.operations = du.filter(proc.operations, args.ignore)
 
     run()
 
@@ -19,7 +19,11 @@ def boot():
 def run():
     ray.init()
 
-    proc.exec_test(cfg.MATRIX_SIZE)
+    data_res = proc.exec_test(cfg.MATRIX_SIZE)
+
+    store_data('test/single', data_res[0])
+    store_data('test/parallel', data_res[1])
+
 
     ray.shutdown()
 

### 2023-07-06 src/mat/matrix.py

@@ -149,12 +149,12 @@ class Matrix:
 
         return submatrices
 
-    # def rank(self):
-    #     j1 = min(self.size()['rows'], self.size()['columns'])
+    def rank(self):
+        j1 = min(self.size()['rows'], self.size()['columns'])
 
-    #     for i in range(j1, 1, -1):
-    #         if bm.get_submatrices_rank(self, i) != -1:
-    #             return i
+        for i in range(j1, 1, -1):
+            if bm.get_submatrices_rank(self, i) != -1:
+                return i
 
 
 

### 2023-07-06 src/mat/raymatrix.py

@@ -145,10 +145,9 @@ class RayMatrix(Matrix):
     def task_sum(results):
         return sum(results)
 
-    @staticmethod
-    def product(a, b):
-        a_columns = a.size()["columns"]
-        a_rows = a.size()["rows"]
+    def product(self, b):
+        a_columns = self.size()["columns"]
+        a_rows = self.size()["rows"]
 
         b_columns = b.size()["columns"]
 
@@ -156,7 +155,7 @@ class RayMatrix(Matrix):
             raise ValueError("Number of columns of the first matrix must match the number of rows of the second matrix")
 
         else:
-            a_elements = a.get()
+            a_elements = self.get()
             b_elements = b.get()
 
             result = [[0] * b_columns for _ in range(a_rows)]

### 2023-07-06 src/process/generator.py

@@ -1,11 +1,5 @@
 import random
 
 def generate_matrix(rows, cols):
-    mat = []
-    print(rows, cols)
-    
-    for i in range(rows):
-        for j in range(cols):
-            mat[i][j] = random.randint(0, 99)
-
+    mat = [[random.randint(0, 99) for _ in range(cols)] for _ in range(rows)]
     return mat

### 2023-07-06 src/process/processor.py

@@ -1,8 +1,10 @@
+import random
 from process.generator import generate_matrix
 import utils.config as cfg
 from mat.matrix import Matrix
 from mat.raymatrix import RayMatrix
 from time import time
+
 
 def exec_test(size):
     a = generate_matrix(size, size)
@@ -14,28 +16,74 @@ def exec_test(size):
     b_base = Matrix(b)
     b_ray = RayMatrix(b)
 
-    for op in operations():
-        exec_time_single = op(a_base, b_base)
-        exec_time_parallel = op(a_ray, b_ray)
+    single_out = {}
+    parallel_out = {}
 
-        print('time single: ', exec_time_single)
-        print('time parallel: ', exec_time_parallel)
-    
-def product(a, b):
+    for func, call in operations.items():
+        exec_time_single = call((a_base, b_base))
+        exec_time_parallel = call((a_ray, b_ray))
+
+        single_out[func] = exec_time_single
+        parallel_out[func] = exec_time_parallel
+
+        print('time single - ', func, ':', exec_time_single)
+        print('time parallel - ', func, ':', exec_time_parallel)
+
+    return single_out, parallel_out
+
+
+def product(args):
     start = time()
-    
-    return time() - start()
 
-def determinant(a):
-    pass
+    a = args[0]
+    b = args[1]
 
-def inverse(a):
+    print(a.product(b))
 
-    pass
+    return time() - start
+
+
+def determinant(args):
+    start = time()
+
+    a = args[0]
+
+    print(a.det())
+    return time() - start
+
+
+def inverse(args):
+    start = time()
+
+    a = args[0]
+
+    print(a.inv())
+    return time() - start
+
+
+def rank(args):
+    start = time()
+
+    a = args[0]
+
+    print(a.rank())
+    return time() - start
+
+
+def scalar_product(args):
+    start = time()
+
+    a = args[0]
+    scalar = random.randint(1, 8)
+
+    print(a.scalar_product(scalar))
+    return time() - start
 
 
 operations = {
     "p": product,
     "d": determinant,
-    "i": inverse
+    "i": inverse,
+    "sp": scalar_product,
+    "r": rank
 }

### 2023-07-06 src/startup/cmd.py

@@ -7,7 +7,7 @@ parser = ArgumentParser(
 
 
 def init_parser():
-    parser.add_argument('-i', '--ignore', nargs="*", choices=["p", "d", "i"], help="ignora un'operazione matriciale[p: prodotto, d: determinante, i: inversa]")
+    parser.add_argument('-i', '--ignore', nargs="*", choices=["p", "d", "i", "sp", "r"], help="ignora un'operazione matriciale[p: prodotto, d: determinante, i: inversa, sp: prodotto scalare, r: rango]")
     parser.add_argument('-l', '--load', nargs=1, type=unsigned, required=True)
 
 

### 2023-07-06 src/utils/data_utils.py

@@ -1,0 +1,12 @@
+import os
+
+
+PROJECT_ROOT = os.path.abspath(os.curdir)
+
+
+def to_ppath(path: str) -> str:
+    return os.path.join(PROJECT_ROOT, path)
+
+
+def data_path(filename: str) -> str:
+    return to_ppath("..\\data\\" + filename)

### 2023-07-06 src/utils/dict_utils.py

@@ -3,3 +3,11 @@ def filter(a, keys):
         del a[e]
 
     return a
+
+
+def csv_to_dict(csv: [dict]):
+    for k, e in enumerate(csv):
+        for f, i in e.items():
+            if i.isdigit():
+                e[f] = int(i)
+    return csv

### 2023-07-06 src/main.py
fix

@@ -24,7 +24,6 @@ def run():
     store_data('test/single', data_res[0])
     store_data('test/parallel', data_res[1])
 
-
     ray.shutdown()
 
 

### 2023-07-06 src/raytest.py

@@ -1,23 +1,0 @@
-from time import sleep
-import ray
-
-ray.init()
-
-sleep(10)
-
-@ray.remote
-def cube(n):
-    return n * n * n
-
-@ray.remote
-def factorial(n):
-    tmp = 1
-    
-    for i in range(1, n):
-        tmp *= i
-
-    return tmp
-
-futures = [factorial.remote(i) for i in range(20, 50)]
-
-print(ray.get(futures))

### 2023-07-06 src/main.py
fix

@@ -4,6 +4,8 @@ import ray
 import process.processor as proc
 import utils.dict_utils as du
 from data.storage_manager import store_data
+
+
 def boot():
     init_parser()
     args = get_args()

### 2023-07-06 src/mat/raymatrix.py
fixed get_square_submatrices() for antipattern #1

@@ -41,19 +41,21 @@ class RayMatrix(Matrix):
         rows = self.size()["rows"]
         cols = self.size()["columns"]
         submatrices = []
-
+        futures = []
+        
         for start_row in range(rows - order + 1):
+            
             for start_col in range(cols - order + 1):
-                submatrix = []
+                submatrix_rows = []
                 
                 for row in range(order):
-                    futures = self.task_get_square_submatrix.remote(self=self, start_row=start_row, start_col=start_col, row=row, order=order)
-                    
-                    submatrix.append(ray.get(futures))
+                    futures.append(self.task_get_square_submatrix.remote(self=self, start_row=start_row, start_col=start_col, row=row, order=order))
                 
-                submatrices.append(RayMatrix(submatrix))
+                submatrix_rows = ray.get(futures)
+                submatrices.append(RayMatrix(submatrix_rows))
 
         return submatrices
+
 
     def minor(self, i, j):
         minor_elements = [row[:j] + row[j + 1:] for row_idx, row in enumerate(self.elements) if row_idx != i]

### 2023-07-06 src/test.py

@@ -4,19 +4,6 @@ from mat.matrix import Matrix
 import ray
 import random
 
-
-# def main():
-#     counters = [rm.ParallelMatrix.remote([
-#         [4, 5, 7],
-#         [5, 3, 2],
-#         [9, 4, 6],
-#         [1, 3, 8]
-#     ]) for i in range(4)]
-
-#     futures = [c.range.remote() for c in counters]
-#     print(ray.get(futures))
-
-
 def main():
     ray.init()
     A = RayMatrix([
@@ -52,8 +39,8 @@ def main():
     # print(RayMatrix.product(B,C))
     # print(E.det())
 
-    print(F.det())
-    print(E.det())
+    print(B)
+    [print(m) for m in B.get_square_submatrices(2)]
     ray.shutdown()
 
 

### 2023-07-06 src/mat/raymatrix.py
fixed rank anti pattern

@@ -42,7 +42,7 @@ class RayMatrix(Matrix):
         cols = self.size()["columns"]
         submatrices = []
         futures = []
-        
+
         for start_row in range(rows - order + 1):
             
             for start_col in range(cols - order + 1):
@@ -88,25 +88,23 @@ class RayMatrix(Matrix):
         columns = self.size()["columns"]
 
         j1 = min(rows, columns)
-
-        for j in range(j1, 1, -1):            
+        for j in range(j1, 1, -1):
             jth_submatrices = self.get_square_submatrices(j)
-            
-
             tasks = []
 
             for submatrix in jth_submatrices:
                 tasks.append(self.task_rank_det.remote(submatrix, j))
 
-
-            ready_tasks, _ = ray.wait(tasks, num_returns = max(rows, columns) - j + 1)
+            ready_tasks, _ = ray.wait(tasks, num_returns=max(rows, columns) - j + 1)
+            
             for ready_task in ready_tasks:
-                    result = ray.get(ready_task)
+                result = ray.get(ready_task)
+                
+                if result is not None:
+                    return result
 
-                    if result is not None:
-                        return result
-                    
         return 0
+
     
     @ray.remote
     def task_inv_cof(self, a, i, j):

### 2023-07-06 src/test.py

@@ -39,8 +39,7 @@ def main():
     # print(RayMatrix.product(B,C))
     # print(E.det())
 
-    print(B)
-    [print(m) for m in B.get_square_submatrices(2)]
+    print(F.rank())
     ray.shutdown()
 
 

### 2023-07-06 src/mat/raymatrix.py
fixed get_square_matrices return value

@@ -45,14 +45,15 @@ class RayMatrix(Matrix):
 
         for start_row in range(rows - order + 1):
             
-            for start_col in range(cols - order + 1):
-                submatrix_rows = []
-                
+            for start_col in range(cols - order + 1):                
                 for row in range(order):
                     futures.append(self.task_get_square_submatrix.remote(self=self, start_row=start_row, start_col=start_col, row=row, order=order))
                 
-                submatrix_rows = ray.get(futures)
-                submatrices.append(RayMatrix(submatrix_rows))
+        submatrix_rows = ray.get(futures)
+                
+        for i in range(0, len(submatrix_rows), order):
+            sub_list = submatrix_rows[i:i+order]
+            submatrices.append(RayMatrix(sub_list))
 
         return submatrices
 

### 2023-07-06 src/test.py

@@ -39,7 +39,10 @@ def main():
     # print(RayMatrix.product(B,C))
     # print(E.det())
 
-    print(F.rank())
+    print(B)
+    print(len(B.get_square_submatrices(4)))
+    [print(m) for m in B.get_square_submatrices(4)]
+    B.get_square_submatrices(2)
     ray.shutdown()
 
 

### 2023-07-06 src/mat/matrix.py
get_square_submatrices cleanup

@@ -132,9 +132,6 @@ class Matrix:
         return Matrix(result)
 
     def get_square_submatrices(self, order):
-        '''
-        Returns an array of square submatrices
-        '''
         rows = self.size()["rows"]
         cols = self.size()["columns"]
         submatrices = []

### 2023-07-06 src/mat/raymatrix.py

@@ -10,10 +10,6 @@ class RayMatrix(Matrix):
     def task_rank_det(submatrix, j):
         if submatrix.det() != 0:
             return j
-    
-    @ray.remote
-    def task_get_square_submatrix(self, start_row, start_col, row, order):
-        return self.elements[start_row + row][start_col:start_col + order]
 
     @ray.remote
     def task_det(self, col_index):
@@ -37,6 +33,15 @@ class RayMatrix(Matrix):
 
         return row
 
+    @ray.remote
+    def task_get_square_submatrix(self, start_row, start_col, order):
+        futures = []
+
+        for row in range(order):
+            futures.append(self.get()[start_row + row][start_col:start_col + order])
+
+        return futures
+
     def get_square_submatrices(self, order):
         rows = self.size()["rows"]
         cols = self.size()["columns"]
@@ -44,18 +49,11 @@ class RayMatrix(Matrix):
         futures = []
 
         for start_row in range(rows - order + 1):
-            
-            for start_col in range(cols - order + 1):                
-                for row in range(order):
-                    futures.append(self.task_get_square_submatrix.remote(self=self, start_row=start_row, start_col=start_col, row=row, order=order))
-                
-        submatrix_rows = ray.get(futures)
-                
-        for i in range(0, len(submatrix_rows), order):
-            sub_list = submatrix_rows[i:i+order]
-            submatrices.append(RayMatrix(sub_list))
 
-        return submatrices
+            for start_col in range(cols - order + 1):
+                futures.append(self.task_get_square_submatrix.remote(self=self, start_row=start_row, start_col=start_col, order=order))
+
+        return [RayMatrix(m) for m in ray.get(futures)]
 
 
     def minor(self, i, j):

### 2023-07-06 src/test.py

@@ -1,3 +1,4 @@
+import time
 from mat.raymatrix import RayMatrix
 from mat.matrix import Matrix
 
@@ -5,7 +6,6 @@ import ray
 import random
 
 def main():
-    ray.init()
     A = RayMatrix([
         [1, 2, 3],
         [4, 5, 7],
@@ -39,10 +39,21 @@ def main():
     # print(RayMatrix.product(B,C))
     # print(E.det())
 
-    print(B)
-    print(len(B.get_square_submatrices(4)))
-    [print(m) for m in B.get_square_submatrices(4)]
-    B.get_square_submatrices(2)
+    elements = [[random.randint(0, 99) for _ in range(100)] for _ in range(100)]
+
+    # [print(m) for m in Matrix(elements).get_square_submatrices(2)]
+
+
+    ray.init(num_cpus=12)
+
+    start = time.time()
+    Matrix(elements).get_square_submatrices(2)
+    print("standard duration = ", time.time() - start)
+
+    start = time.time()
+    RayMatrix(elements).get_square_submatrices(2)
+    print("ray duration = ", time.time() - start)
+
     ray.shutdown()
 
 

### 2023-07-06 src/mat/raymatrix.py
performance enhancements

@@ -34,24 +34,25 @@ class RayMatrix(Matrix):
         return row
 
     @ray.remote
-    def task_get_square_submatrix(self, start_row, start_col, order):
+    def task_get_square_submatrix(elements, start_row, start_col, order):
         futures = []
 
         for row in range(order):
-            futures.append(self.get()[start_row + row][start_col:start_col + order])
+            futures.append(elements[start_row + row][start_col:start_col + order])
 
         return futures
 
     def get_square_submatrices(self, order):
+        elements_id = ray.put(self.get())
+
         rows = self.size()["rows"]
         cols = self.size()["columns"]
-        submatrices = []
         futures = []
 
         for start_row in range(rows - order + 1):
 
             for start_col in range(cols - order + 1):
-                futures.append(self.task_get_square_submatrix.remote(self=self, start_row=start_row, start_col=start_col, order=order))
+                futures.append(self.task_get_square_submatrix.remote(elements=elements_id, start_row=start_row, start_col=start_col, order=order))
 
         return [RayMatrix(m) for m in ray.get(futures)]
 

### 2023-08-08 scripts/plot.py
project reformat

@@ -1,38 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import csv
-import math
-
-from src.data.storage_manager import load_data
-
-mongo = []
-neo = []
-queryn = 4
-
-for i in range(25, 101, 25):
-    single_data = load_data('single')
-    parallel_data = load_data('parallel')
-
-
-labels = ['25%', '50%', '75%', '100%']
-x = np.arange(len(labels))  # the label locations
-width = 0.35  # the width of the bars
-
-fig, ax = plt.subplots()
-
-rects1 = ax.bar(x - width / 2, neo, width, label='Neo4j')
-rects2 = ax.bar(x + width / 2, mongo, width, label='MongoDB')
-
-# Add some text for labels, title and custom x-axis tick labels, etc.
-
-ax.set_ylabel('Tempo (Millisecondi)')
-ax.set_title('3 WHERE 2 JOIN')
-ax.set_xticks(x, labels)
-ax.legend()
-
-ax.bar_label(rects1, padding=3)
-ax.bar_label(rects2, padding=3)
-
-fig.tight_layout()
-
-plt.show()

### 2023-08-08 src/data/storage_manager.py

@@ -1,21 +1,0 @@
-from utils.data_utils import data_path
-from utils.dict_utils import csv_to_dict
-import csv
-
-
-def store_data(filename, data_model):
-    with open(data_path(filename + '.csv'), "w", newline="") as fq:
-        writer = csv.writer(fq)
-        writer.writerow(['Run type:', filename.capitalize()])
-
-        for op, time in data_model.items():
-            writer.writerow([op, time])
-
-        fq.close()
-
-
-def load_data(filename):
-    with open(data_path(filename + '.csv'), 'w', newline="") as fq:
-        data = csv_to_dict(csv_to_dict(list(csv.DictReader(fq))))
-
-        return data

### 2023-08-08 src/main.py

@@ -1,33 +1,0 @@
-from startup.cmd import get_args, init_parser
-import utils.config as cfg
-import ray
-import process.processor as proc
-import utils.dict_utils as du
-from data.storage_manager import store_data
-
-
-def boot():
-    init_parser()
-    args = get_args()
-
-    cfg.MATRIX_SIZE = int(args.load[0])
-
-    if args.ignore is not None:
-        proc.operations = du.filter(proc.operations, args.ignore)
-
-    run()
-
-
-def run():
-    ray.init()
-
-    data_res = proc.exec_test(cfg.MATRIX_SIZE)
-
-    store_data('test/single', data_res[0])
-    store_data('test/parallel', data_res[1])
-
-    ray.shutdown()
-
-
-if __name__ == "__main__":
-    boot()

### 2023-08-08 src/module/__init__.py

@@ -1,0 +1,10 @@
+import ray
+
+def run():
+    ray.init()
+
+    ray.shutdown()
+
+
+if __name__ == "__main__":
+    run()

### 2023-08-08 src/module/classes/matrix.py

@@ -88,24 +88,23 @@ class Matrix:
 
     def det(self):
         if self.is_square():
-            size = self.size()["rows"]
-            a = self.get()
+            cols = self.size()["columns"]
+            elements = self.get()
 
-            if size == 1:
-                return a[0][0]
+            if cols == 1:
+                return elements[0][0]
 
-            elif size == 2:
-                return (a[0][0] * a[1][1]) - (a[0][1] * a[1][0])
+            elif cols == 2:
+                return (elements[0][0] * elements[1][1]) - (elements[0][1] * elements[1][0])
 
             else:
+                submatrices = [self.minor(0, j) for j in range(cols)]
+                results = [submatrix.det() for submatrix in submatrices]
 
-                for i in range(size):
-                    sum = 0
+                return sum(results)
 
-                    for j in range(size):
-                        sum += a[i][j] * ((-1) ** (i + j + 2)) * self.minor(i, j).det()
-
-                return sum
+        else:
+            raise ValueError("Cannot compute determinant of a non-square matrix")
 
     def product(self, b):
         a_columns = self.size()["columns"]

### 2023-08-08 src/module/classes/raymatrix.py

@@ -11,20 +11,6 @@ class RayMatrix(Matrix):
         if submatrix.det() != 0:
             return j
 
-    @ray.remote
-    def task_det(self, col_index):
-        elems = self.get()
-        submat = self.minor(0, col_index)
-
-        submat_det_sum = 0
-
-        for i in range(len(elems)):
-            submatrix_det = submat.det()
-
-            submat_det_sum += elems[i][col_index] * ((-1) ** (i + col_index + 2)) * submatrix_det
-
-        return submat_det_sum
-
     @staticmethod
     @ray.remote
     def task_scalar_product(row, scalar):
@@ -33,55 +19,81 @@ class RayMatrix(Matrix):
 
         return row
 
-    @ray.remote
-    def task_get_square_submatrix(elements, start_row, start_col, order):
-        futures = []
+    # @ray.remote
+    # def task_gen_submatrix(elements, start_row, start_col, order):
+    #     futures = []
 
-        for row in range(order):
-            futures.append(elements[start_row + row][start_col:start_col + order])
+    #     for row in range(order):
+    #         futures.append(elements[start_row + row][start_col:start_col + order])
 
-        return futures
+    #     return futures
 
-    def get_square_submatrices(self, order):
-        elements_id = ray.put(self.get())
+    # def get_square_submatrices(self, order):
+    #     elements_id = ray.put(self.get())
 
-        rows = self.size()["rows"]
-        cols = self.size()["columns"]
-        futures = []
+    #     rows = self.size()["rows"]
+    #     cols = self.size()["columns"]
+    #     futures = []
 
-        for start_row in range(rows - order + 1):
+    #     for start_row in range(rows - order + 1):
+    #         for start_col in range(cols - order + 1):
+    #             futures.append(self.task_gen_submatrix.remote(elements=elements_id, start_row=start_row, start_col=start_col, order=order))
 
-            for start_col in range(cols - order + 1):
-                futures.append(self.task_get_square_submatrix.remote(elements=elements_id, start_row=start_row, start_col=start_col, order=order))
-
-        return [RayMatrix(m) for m in ray.get(futures)]
+    #     return [RayMatrix(m) for m in ray.get(futures)]
 
 
     def minor(self, i, j):
         minor_elements = [row[:j] + row[j + 1:] for row_idx, row in enumerate(self.elements) if row_idx != i]
         return RayMatrix(minor_elements)
 
+    @staticmethod
+    @ray.remote
+    def task_det(submatrix):
[…]

### 2023-08-08 src/process/generator.py

@@ -1,5 +1,0 @@
-import random
-
-def generate_matrix(rows, cols):
-    mat = [[random.randint(0, 99) for _ in range(cols)] for _ in range(rows)]
-    return mat

### 2023-08-08 src/process/processor.py

@@ -1,89 +1,0 @@
-import random
-from process.generator import generate_matrix
-import utils.config as cfg
-from mat.matrix import Matrix
-from mat.raymatrix import RayMatrix
-from time import time
-
-
-def exec_test(size):
-    a = generate_matrix(size, size)
-    b = generate_matrix(size, size)
-    
-    a_base = Matrix(a)
-    a_ray = RayMatrix(a)
-
-    b_base = Matrix(b)
-    b_ray = RayMatrix(b)
-
-    single_out = {}
-    parallel_out = {}
-
-    for func, call in operations.items():
-        exec_time_single = call((a_base, b_base))
-        exec_time_parallel = call((a_ray, b_ray))
-
-        single_out[func] = exec_time_single
-        parallel_out[func] = exec_time_parallel
-
-        print('time single - ', func, ':', exec_time_single)
-        print('time parallel - ', func, ':', exec_time_parallel)
-
-    return single_out, parallel_out
-
-
-def product(args):
-    start = time()
-
-    a = args[0]
-    b = args[1]
-
-    print(a.product(b))
-
-    return time() - start
-
-
-def determinant(args):
-    start = time()
-
-    a = args[0]
-
-    print(a.det())
-    return time() - start
-
-
-def inverse(args):
-    start = time()
-
-    a = args[0]
-
-    print(a.inv())
-    return time() - start
-
-
-def rank(args):
-    start = time()
-
-    a = args[0]
-
-    print(a.rank())
-    return time() - start
-
-
-def scalar_product(args):
-    start = time()
-
-    a = args[0]
-    scalar = random.randint(1, 8)
-
-    print(a.scalar_product(scalar))
-    return time() - start
-
-
-operations = {
-    "p": product,
-    "d": determinant,
-    "i": inverse,
-    "sp": scalar_product,
-    "r": rank
-}

### 2023-08-08 src/startup/cmd.py

@@ -1,15 +1,0 @@
-from argparse import ArgumentParser
-from startup.cmdtype import unsigned
-
-parser = ArgumentParser(
-    prog="Progetto di Laboratorio di reti e sistemi distribuiti",
-    description="Progetto creato per Laboratorio di reti e sistemi distribuiti")
-
-
-def init_parser():
-    parser.add_argument('-i', '--ignore', nargs="*", choices=["p", "d", "i", "sp", "r"], help="ignora un'operazione matriciale[p: prodotto, d: determinante, i: inversa, sp: prodotto scalare, r: rango]")
-    parser.add_argument('-l', '--load', nargs=1, type=unsigned, required=True)
-
-
-def get_args():
-    return parser.parse_args()

### 2023-08-08 src/startup/cmdtype.py

@@ -1,8 +1,0 @@
-from argparse import ArgumentTypeError
-
-
-def unsigned(x): 
-    x = int(x)
-    if x < 0:
-        raise ArgumentTypeError("Il valore inserito deve essere > 0")
-    return x

### 2023-08-08 src/test.py

@@ -1,6 +1,6 @@
 import time
-from mat.raymatrix import RayMatrix
-from mat.matrix import Matrix
+from module.classes.raymatrix import RayMatrix
+from module.classes.matrix import Matrix
 
 import ray
 import random
@@ -39,19 +39,19 @@ def main():
     # print(RayMatrix.product(B,C))
     # print(E.det())
 
-    elements = [[random.randint(0, 99) for _ in range(100)] for _ in range(100)]
+    elements = [[random.randint(0, 99) for _ in range(10)] for _ in range(10)]
 
     # [print(m) for m in Matrix(elements).get_square_submatrices(2)]
 
 
-    ray.init(num_cpus=12)
+    ray.init(num_cpus=12, ignore_reinit_error=True)
 
     start = time.time()
-    Matrix(elements).get_square_submatrices(2)
+    print(Matrix(elements).det())
     print("standard duration = ", time.time() - start)
 
     start = time.time()
-    RayMatrix(elements).get_square_submatrices(2)
+    print(RayMatrix(elements).det())
     print("ray duration = ", time.time() - start)
 
     ray.shutdown()

### 2023-08-08 src/utils/config.py

@@ -1,1 +1,0 @@
-MATRIX_SIZE = 0

### 2023-08-08 src/utils/data_utils.py

@@ -1,12 +1,0 @@
-import os
-
-
-PROJECT_ROOT = os.path.abspath(os.curdir)
-
-
-def to_ppath(path: str) -> str:
-    return os.path.join(PROJECT_ROOT, path)
-
-
-def data_path(filename: str) -> str:
-    return to_ppath("..\\data\\" + filename)

### 2023-08-08 src/utils/dict_utils.py

@@ -1,13 +1,0 @@
-def filter(a, keys):
-    for e in keys:
-        del a[e]
-
-    return a
-
-
-def csv_to_dict(csv: [dict]):
-    for k, e in enumerate(csv):
-        for f, i in e.items():
-            if i.isdigit():
-                e[f] = int(i)
-    return csv

### 2023-08-08 distmat/__init__.py
Matrix class reformat

@@ -1,10 +1,8 @@
 import ray
-
-def run():
-    ray.init()
-
-    ray.shutdown()
-
+import logging
 
 if __name__ == "__main__":
-    run()
+    # if ray.is_initialized:
+    #     ray.shutdown()
+    # ray.init(logging_level=logging.ERROR)
+    

### 2023-08-08 distmat/classes/matrix.py

@@ -1,158 +1,0 @@
-from mat import base_ops as bm
-
-
-class Matrix:
-    def __init__(self, elements):
-        for i in elements:
-            if len(i) != len(elements[0]):
-                raise ValueError("Invalid row size")
-
-        self.elements = elements
-
-    def __str__(self):
-        tmp = "\n"
-
-        max_lengths = []
-        for i in range(len(self.elements[0])):
-            column_lengths = [len(str(row[i])) for row in self.elements]
-            max_lengths.append(max(column_lengths))
-
-        for row in self.elements:
-            tmp += "| "
-            for i in range(len(self.elements[0])):
-                tmp += f"{str(row[i]):^{max_lengths[i]}} "
-            tmp += "|\n"
-
-        return tmp
-
-    def size(self):
-        rows = len(self.elements)
-        columns = len(self.elements[0])
-
-        return {"rows": rows, "columns": columns}
-
-    def get(self):
-        return self.elements
-
-    def transpose(self):
-        a = self.get()
-        rows = self.size()["rows"]
-        columns = self.size()["columns"]
-
-        transpose_matrix = [[0] * rows for _ in range(columns)]
-
-        for i in range(rows):
-            for j in range(columns):
-                transpose_matrix[j][i] = a[i][j]
-
-        return Matrix(transpose_matrix)
-
-    def is_square(self):
-        return self.size()["rows"] == self.size()["columns"]
-
-    def minor(self, i, j):
-        minor_elements = [row[:j] + row[j + 1:] for row_idx, row in enumerate(self.elements) if row_idx != i]
-        return Matrix(minor_elements)
-
-    def scalar_product(self, scalar):
-        elems = self.get()
-
-        rows = self.size()['rows']
-        cols = self.size()['columns']
-
-        for i in range(rows):
-            for j in range(cols):
-                elems[i][j] *= scalar
-
-        return Matrix(elems)
-
-    def inv(self):
-        if not self.is_square():
-            raise Exception("Matrix must be square")
-
-        elif Matrix.det(self) == 0:
-            raise Exception("Matrix is not invertible")
-
-        else:
-            size = self.size()["rows"]
-            a = self.get()
-            cof_arr = [[0] * size for _ in range(size)]
-
-            for i in range(size):
-                for j in range(size):
-                    cof_arr[i][j] = a[i][j] * ((-1) ** (i + j + 2)) * self.minor(i, j).det()
-
[…]

### 2023-08-08 distmat/matrix.py

@@ -1,0 +1,180 @@
+
+class Matrix:
+    def __init__(self, elements):
+        for i in elements:
+            if len(i) != len(elements[0]):
+                raise ValueError("Invalid row size")
+
+        self.elements = elements
+
+        self.rows = len(self.elements[0])
+        self.columns = len(self.elements)
+
+    def __str__(self):
+        tmp = "\n"
+
+        rows, _ = self.get_size()
+        elements = self.get_elements()
+
+        max_lengths = []
+        for i in range(rows):
+            column_lengths = [len(str(row[i])) for row in elements]
+            max_lengths.append(max(column_lengths))
+
+        for row in elements:
+            tmp += "| "
+            for i in range(rows):
+                tmp += f"{str(row[i]):^{max_lengths[i]}} "
+            tmp += "|\n"
+
+        return tmp
+
+    def get_elements(self):
+        '''
+        Returns a 2D array of the matrix elements 
+        '''
+
+        return self.elements
+
+    def get_size(self):
+        '''
+        Returns the tuple (numberof_rows, numberof_columns)
+        '''
+
+        return (self.rows, self.columns)
+
+    def transpose(self):
+        elements = self.get_elements()
+        rows, columns = self.get_size()
+
+        transpose_matrix = [[0] * rows for _ in range(columns)]
+
+        for i in range(rows):
+            for j in range(columns):
+                transpose_matrix[j][i] = elements[i][j]
+
+        return Matrix(transpose_matrix)
+
+    def is_square(self):
+        rows, columns = self.get_size()
+
+        return rows == columns
+
+    def minor(self, i, j):
+        '''
+        Extract a minor matrix by removing the ith row and jth column
+        '''
+
+        elements = self.get_elements()
+        minor_elements = [row[:j] + row[j + 1:] for row_idx, row in enumerate(elements) if row_idx != i]
+
+        return Matrix(minor_elements)
+
+    def scalar_product(self, scalar):
+        '''
+        Multiply a matrix by a scalar
+        '''
+
+        elements = self.get_elements()
+        rows, columns = self.get_size()
+
+        for i in range(rows):
+            for j in range(columns):
+                elements[i][j] *= scalar
+
+        return Matrix(elements)
+
+    def inv(self):
+        if not self.is_square():
+            raise Exception("Matrix must be square")
+
+        elif Matrix.det(self) == 0:
+            raise Exception("Matrix is not invertible")
+
+        else:
+            elements = self.get_elements()
+            rows, _ = self.get_size()
+
[…]

### 2023-08-08 distmat/raymatrix.py

@@ -1,7 +1,8 @@
-from mat.matrix import Matrix
+from . import Matrix
 import ray
 
 
+@ray.remote
 class RayMatrix(Matrix):
     def __init__(self, elements):
         super().__init__(elements)
@@ -63,23 +64,10 @@ class RayMatrix(Matrix):
                 return (elements[0][0] * elements[1][1]) - (elements[0][1] * elements[1][0])
 
             else:
-                # Iterative computation of determinant
-                submatrices = [[self.minor(row, col) for col in range(cols)] for row in range(cols)]
-                determinants = []
-                batch_size = 10  # Adjust batch size as needed
+                submatrices = [self.minor(0, j) for j in range(cols)]
+                results = [submatrix.det() for submatrix in submatrices]
 
-                for i in range(0, cols, batch_size):
-                    batch_results = ray.get([self.task_det.remote(submatrix) for submatrix_row in submatrices[i:i+batch_size] for submatrix in submatrix_row])
-                    determinants.extend(batch_results)
-
-                determinant = 0
-
-                for col in range(cols):
-                    sign = (-1) ** col
-                    sub_determinants = [determinants[row * cols + col] for row in range(cols)]
-                    determinant += sign * elements[0][col] * self._cofactor(sub_determinants)
-
-                return determinant
+                return sum(results)
 
         else:
             raise ValueError("Cannot compute determinant of a non-square matrix")

### 2023-08-08 distmat/scripts/test.py

@@ -1,62 +1,0 @@
-import time
-from module.classes.raymatrix import RayMatrix
-from module.classes.matrix import Matrix
-
-import ray
-import random
-
-def main():
-    A = RayMatrix([
-        [1, 2, 3],
-        [4, 5, 7],
-        [5, 3, 2],
-        [9, 5, 6],
-        [1, 3, 8]])
-
-    B = RayMatrix([[random.randint(0, 99) for _ in range(5)] for _ in range(5)])
-    C = RayMatrix([[random.randint(0, 99) for _ in range(40)] for _ in range(40)])
-
-    D = RayMatrix([[1, 2, 3],
-                   [4, 5, 6],
-                   [7, 8, 9]])
-
-    E = RayMatrix([
-        [4, 10, 2, 6],
-        [8, 9, 1, 3],
-        [5, 8, 4, 1],
-        [6, 3, 0, 2]
-    ])
-    F = Matrix([
-        [4, 10, 2, 6],
-        [8, 9, 1, 3],
-        [5, 8, 4, 1],
-        [6, 3, 0, 2]
-    ])
-
-    # print(E.inv())
-    # print(F.inv())
-
-    # print(RayMatrix.product(B,C))
-    # print(E.det())
-
-    elements = [[random.randint(0, 99) for _ in range(10)] for _ in range(10)]
-
-    # [print(m) for m in Matrix(elements).get_square_submatrices(2)]
-
-
-    ray.init(num_cpus=12, ignore_reinit_error=True)
-
-    start = time.time()
-    print(Matrix(elements).det())
-    print("standard duration = ", time.time() - start)
-
-    start = time.time()
-    print(RayMatrix(elements).det())
-    print("ray duration = ", time.time() - start)
-
-    ray.shutdown()
-
-
-if __name__ == "__main__":
-    main()
-

### 2023-08-08 distmat/test.py

@@ -1,0 +1,21 @@
+
+from matrix import Matrix
+
+import random
+
+def main():
+    A = Matrix([
+        [4, 0, 2, 6],
+        [8, 9, 1, 3],
+        [5, 8, 4, 1],
+        [6, 3, 0, 2]
+    ])
+
+    B = Matrix([[random.randint(0, 99) for _ in range(4)] for _ in range(4)])
+
+    print(A, B)
+    print(Matrix.product(A,B))
+
+if __name__ == "__main__":
+    main()
+

### 2023-08-08 distmat/matrix.py
fixed det

@@ -5,7 +5,7 @@ class Matrix:
             if len(i) != len(elements[0]):
                 raise ValueError("Invalid row size")
 
-        self.elements = elements
+        self.elements = [[float(val) for val in row] for row in elements]
 
         self.rows = len(self.elements[0])
         self.columns = len(self.elements)
@@ -43,6 +43,34 @@ class Matrix:
 
         return (self.rows, self.columns)
 
+    def is_square(self):
+        rows, columns = self.get_size()
+
+        return rows == columns
+
+    @staticmethod
+    def product(A, B):
+        a_rows, a_columns = A.get_size()
+        _, b_columns = B.get_size()
+
+        if a_rows != b_columns:
+            raise ValueError("Number of columns of first matrix must match the number of rows of second matrix")
+
+        else:
+            a_elements = A.get_elements()
+            b_elements = B.get_elements()
+
+            result = [[[0] * i for i in range(b_columns)] for j in range(a_rows)]
+
+            for i in range(a_rows):
+                for j in range(b_columns):
+                    result[i][j] = 0
+
+                    for k in range(a_columns):
+                        result[i][j] += a_elements[i][k] * b_elements[k][j]
+
+        return Matrix(result)
+    
     def transpose(self):
         elements = self.get_elements()
         rows, columns = self.get_size()
@@ -55,10 +83,6 @@ class Matrix:
 
         return Matrix(transpose_matrix)
 
-    def is_square(self):
-        rows, columns = self.get_size()
-
-        return rows == columns
 
     def minor(self, i, j):
         '''
@@ -74,15 +98,16 @@ class Matrix:
         '''
         Multiply a matrix by a scalar
         '''
+        rows, columns = self.get_size()
 
         elements = self.get_elements()
-        rows, columns = self.get_size()
+        result_elements = [[0] * rows for _ in range(columns)]
 
         for i in range(rows):
             for j in range(columns):
-                elements[i][j] *= scalar
+                result_elements[i][j] = elements[i][j] * scalar
 
-        return Matrix(elements)
+        return Matrix(result_elements)
 
     def inv(self):
         if not self.is_square():
@@ -117,40 +142,20 @@ class Matrix:
                 return (elements[0][0] * elements[1][1]) - (elements[0][1] * elements[1][0])
 
             else:
-                minors = [self.minor(0, j) for j in range(columns)]
-                results = [minor.det() for minor in minors]
+                det_value = 0
 
-                return sum(results)
[…]

### 2023-08-08 distmat/test.py

@@ -1,6 +1,6 @@
 
 from matrix import Matrix
-
+import numpy as np
 import random
 
 def main():
@@ -11,10 +11,12 @@ def main():
         [6, 3, 0, 2]
     ])
 
-    B = Matrix([[random.randint(0, 99) for _ in range(4)] for _ in range(4)])
+    elements = [[random.randint(0, 99) for _ in range(8)] for _ in range(8)]
+    B = Matrix(elements)
 
-    print(A, B)
-    print(Matrix.product(A,B))
+    print(np.linalg.det(np.array(elements)))
+    
+    print(B.det())
 
 if __name__ == "__main__":
     main()

### 2023-08-08 distmat/matrix.py
is_vector is_scalar

@@ -5,7 +5,7 @@ class Matrix:
             if len(i) != len(elements[0]):
                 raise ValueError("Invalid row size")
 
-        self.elements = [[float(val) for val in row] for row in elements]
+        self.elements = elements
 
         self.rows = len(self.elements[0])
         self.columns = len(self.elements)
@@ -47,6 +47,15 @@ class Matrix:
         rows, columns = self.get_size()
 
         return rows == columns
+
+    def is_vector(self):
+        rows, columns = self.get_size()
+
+        return (rows == 1 and columns > 1) or (columns == 1 and rows > 1)
+
+    def is_scalar(self):
+        rows, columns = self.get_size()
+        return rows == 1 and columns == 1
 
     @staticmethod
     def product(A, B):
@@ -94,9 +103,9 @@ class Matrix:
 
         return Matrix(minor_elements)
 
-    def scalar_product(self, scalar):
+    def dot(self, scalar):
         '''
-        Multiply a matrix by a scalar
+        
         '''
         rows, columns = self.get_size()
 
@@ -113,22 +122,23 @@ class Matrix:
         if not self.is_square():
             raise Exception("Matrix must be square")
 
-        elif Matrix.det(self) == 0:
+        elif self.det() == 0:
             raise Exception("Matrix is not invertible")
 
         else:
-            elements = self.get_elements()
             rows, _ = self.get_size()
 
             cof_elements = [[0] * rows for _ in range(rows)]
 
             for i in range(rows):
                 for j in range(rows):
-                    cof_elements[i][j] = elements[i][j] * ((-1) ** (i + j + 2)) * self.minor(i, j).det()
+                    cof_elements[i][j] = ((-1) ** (i + j + 2)) * self.minor(i, j).det()
 
             cof_matrix = Matrix(cof_elements)
+            det_reciprocal = 1 / self.det()
+            inv_matrix = cof_matrix.dot(det_reciprocal)
 
-            return cof_matrix.scalar_product(1 / self.det())
+            return inv_matrix
 
     def det(self):
         if self.is_square():

### 2023-08-08 distmat/test.py

@@ -11,12 +11,19 @@ def main():
         [6, 3, 0, 2]
     ])
 
-    elements = [[random.randint(0, 99) for _ in range(8)] for _ in range(8)]
+    elements = [[random.randint(0, 99) for _ in range(1)] for _ in range(1)]
     B = Matrix(elements)
 
-    print(np.linalg.det(np.array(elements)))
-    
-    print(B.det())
+    # print(np.linalg.det(np.array(elements)))
+    # print(B.det())
+
+    # print(Matrix(np.linalg.inv(np.array(elements))))
+    # print(B.inv().transpose())
+
+    # print(Matrix(np.dot(np.array(elements))))
+    # print(B.inv().transpose())
+
+    print(B.is_scalar())
 
 if __name__ == "__main__":
     main()

### 2023-08-08 distmat/matrix.py
get_type

@@ -48,14 +48,21 @@ class Matrix:
 
         return rows == columns
 
-    def is_vector(self):
+    def get_type(self):
         rows, columns = self.get_size()
 
-        return (rows == 1 and columns > 1) or (columns == 1 and rows > 1)
+        if rows == 1 and columns == 1:
+            return 1 # scalar
+        
+        elif rows == 1 and columns > 1:
+            return 2 # row vector
+        
+        elif columns == 1 and rows > 1:
+            return 3 # column vector
+        
+        else:
+            return 4 # matrix
 
-    def is_scalar(self):
-        rows, columns = self.get_size()
-        return rows == 1 and columns == 1
 
     @staticmethod
     def product(A, B):
@@ -103,14 +110,40 @@ class Matrix:
 
         return Matrix(minor_elements)
 
-    def dot(self, scalar):
+    @staticmethod
+    def eq_size(A, B):
+        '''
+        Return true if A and B have the same size
+        '''
+        a_rows, a_columns = A.get_size()
+        b_rows, b_columns = B.get_size()
+
+        if a_rows == b_rows and a_columns == b_columns:
+            return True
+
+        # Check if A and B are both column vectors or row vectors and have the same size
+        if (a_rows == 1 and b_rows == 1) or (a_columns == 1 and b_columns == 1):
+            return True
+
+        return False
+
+
+
+    @staticmethod
+    def dot(A, B):
         '''
         
         '''
-        rows, columns = self.get_size()
+        a_rows, a_columns = A.get_size()
+        b_rows, b_columns = B.get_size()
 
-        elements = self.get_elements()
-        result_elements = [[0] * rows for _ in range(columns)]
+        a_elements = A.get_elements()
+        
+        # result_elements = [[0] * rows for _ in range(columns)]
+
+        if A.
+        if A.is_vector() and B.is_vector():
+            for i in 
 
         for i in range(rows):
             for j in range(columns):

### 2023-09-19 latex/paper/main.tex
report init

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% University Assignment Title Page 
% LaTeX Template
% Version 1.0 (27/12/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% WikiBooks (http://en.wikibooks.org/wiki/LaTeX/Title_Creation)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
% 
% Instructions for using this template:
% This title page is capable of being compiled as is. This is not useful for 
% including it in another document. To do this, you have two options: 
%
% 1) Copy/paste everything between \begin{document} and \end{document} 
% starting at \begin{titlepage} and paste this into another LaTeX file where you 
% want your title page.
% OR
% 2) Remove everything outside the \begin{titlepage} and \end{titlepage} and 
% move this file to the same directory as the LaTeX file you wish to add it to. 
% Then add \input{./title_page_1.tex} to your LaTeX file where you want your
% title page.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\title{Title page with logo}
%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{article}
\usepackage[italian]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{filecontents}
\usepackage{verbatim}
\usepackage{eurosym}
\usepackage[export]{adjustbox}

\usepackage{listings}
\usepackage{xcolor}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}

\begin{document}
\begin{titlepage}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % Defines a new command for the horizontal lines, change thickness here

\center % Center everything on the page

%----------------------------------------------------------------------------------------
%	HEADING SECTIONS
%----------------------------------------------------------------------------------------

\textsc{\LARGE Università di Messina}\\[1.5cm] % Name of your university/college
\textsc{\Large Dipartimento di scienze matematiche e informatiche, scienze fisiche e della terra}\\[0.5cm] % Major heading such as course name
\textsc{\large Corso di Laurea Triennale in Informatica}\\[0.5cm] % Minor heading such as course title

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\HRule \\[0.4cm]
{ \huge \bfseries Distributed computation of linear algebra operations}\\[0.4cm] % Title of your document
\HRule \\[1.5cm]

%----------------------------------------------------------------------------------------
%	AUTHOR SECTION
%----------------------------------------------------------------------------------------

\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\
Gabriele \textsc{Aloisio} \textit{(503264)} \\
\end{flushleft}
\end{minipage}
~
\begin{minipage}{0.4\textwidth}
\begin{flushright} \large
\emph{Supervisor:} \\
prof. Roberto \textsc{Marino} \\
\end{flushright}
\end{minipage}\\[2cm]

% If you don't want a supervisor, uncomment the two lines below and remove the section above
%\Large \emph{Author:}\\
%John \textsc{Smith}\\[3cm] % Your name

%----------------------------------------------------------------------------------------
%	DATE SECTION
%----------------------------------------------------------------------------------------

{\large \today}\\[2cm] % Date, change the \today to a set date if you want to be precise

%----------------------------------------------------------------------------------------
%	LOGO SECTION
%----------------------------------------------------------------------------------------

\includegraphics[width=70px, keepaspectratio]{unime.png}\\[1cm] % Include a department/university logo - this will require the graphicx package

%----------------------------------------------------------------------------------------

\vfill % Fill the rest of the page with whitespace

\end{titlepage}

\tableofcontents
\pagebreak

\section{Abstract}

\section{Introduzione}
Matrix operations are fundamental in numerous scientific and computational domains, serving as the building blocks for various applications. However, traditional sequential computation on a single machine often becomes a bottleneck, limiting the speed and scalability of matrix calculations. To address this problem, we are going to use an open-source distributed computing framework called \textbf{Ray}.

\subsection{Overview of distributed systems and networks}
A distributed system consists of multiple interconnected computers that collaborate and coordinate their activities to achieve a common goal. These systems are designed to tackle tasks that cannot be efficiently computed by a single machine. Networks serve as the backbone of distributed systems, enabling communication among the connected nodes. Network infrastructures enable coordination, data sharing, and synchronization across distributed systems, regardless of their physical locations.

\pagebreak
\subsection{The Ray library}
The Ray library is an open-source distributed computing framework primarily designed for building scalable and high-performance applications. It was developed by the company Anyscale, which aimed to simplify the development of distributed and parallel computing applications. When it comes to matrix operation computation, using Ray for distributed computation offers several advantages over performing the operations on a single machine:
    \begin{itemize}
        \item \textbf{Faster execution}: With Ray you can use multiple machines and CPUs to compute operations. This significantly reduces the computation time compared to a single machine. Each machine can work on a subset of the matrix data, completing the calculations in parallel. This distributed approach can lead to substantial speedups.
        \item \textbf{Scalability}: As the size of the matrices grows, a single machine may struggle to handle the computational demands due to memory limitations or processing power constraints. Ray allows you to scale horizontally by adding more machines to the distributed setup.
        \item \textbf{Fault tolerance}: Ray offers fault tolerance mechanisms that ensure the continuity of computation even in the presence of failures. If a machine participating in the distributed computation fails, Ray can automatically redistribute the workload to other available machines.
        \item \textbf{Resource utilization}: With Ray, each machine contributes its processing power and memory capacity to the overall computation. This efficient utilization of resources allows you to make the most of the available hardware infrastructure, compared to a single machine that may be underutilized.
    \end{itemize}

\pagebreak
\section{Caso di studio}

\section{Implementation}
    \subsection{Python}
    Python is a versatile and high-level programming language known for its simplicity and readability. It is widely used for a variety of applications, including web development, data analysis, artificial intelligence, scientific computing, and automation. It supports multiple programming paradigms, including procedural, object-oriented, and functional programming, and its large and active community contributes to a vast ecosystem of libraries and frameworks, making it a powerful tool for a wide range of tasks. It is one of the two languages supported by Ray, alongside Java.

    \subsection{Ray}
    To parallelize a function with Ray, import Ray and and initialize it with \textit{ray.init()}. Then decorate the function with \textit{@ray.remote} to declare that you want to run this function remotely. Lastly, call the function with \textit{.remote()} instead of calling it normally. This remote call yields a future, a Ray object reference, that you can then fetch with \textit{ray.get}:

    \begin{lstlisting}[language=Python]
import ray
ray.init()

@ray.remote
def f(x):
    return x * x

futures = [f.remote(i) for i in range(4)]
print(ray.get(futures)) # [0, 1, 4, 9]
    \end{lstlisting}

\subsection{Ray clusters}

\pagebreak
\section{Result of tests}

\section{Conclusion}

\end{document}

### 2023-09-19 latex/paper/main.tex
dot product paper

@@
 
+\usepackage[margin=3cm]{geometry}
 \usepackage{listings}
@@
 
-\section{Introduzione}
+\section{Introduction}
 Matrix operations are fundamental in numerous scientific and computational domains, serving as the building blocks for various applications.
@@
 \pagebreak
-\section{Caso di studio}
+\section{Case study}
 
@@
 \pagebreak
+\subsection{The Matrix class}
+Before implementing the parallelized algorithms for Ray, we write the code for the serial execution in the \textbf{Matrix} class (most of the functions defined in the class are auxiliary therefore were excluded for practical purposes):
+\begin{lstlisting}[language=Python, label=Matrix class]
+class Matrix:
+    def __init__(self, data):
+        if isinstance(data, list):
+            self.data = [[round(val, 8) for val in row]
+                                        for row in data]
+        else:
+            raise ValueError("Input must be a list")
+
+    [...]
+
+    def dot(A, B): [...]
+    def det(self): [...]
+    def rank(self): [...]
+    def inv(self): [...]
+\end{lstlisting}
+Which we modify later to achieve parallelization with Ray.
+
+\pagebreak
+\subsubsection{Dot product}
+\textbf{Matrix multiplication} is a binary operation that produces a matrix from two matrices.
+For matrix multiplication, the number of columns in the first matrix must be equal to the number of rows in the second matrix.
+The resulting matrix, known as the matrix \textbf{product}, has the number of rows of the first and the number of columns of the second matrix.
+$$
+    A = \begin{bmatrix}
+        a_{11} & a_{12} & \dots  & a_{1n} \\
+        a_{21} & a_{22} & \dots  & a_{2n} \\
+        \vdots & \vdots & \ddots & \vdots \\
+        a_{m1} & a_{m2} & \dots  & a_{mn}
+    \end{bmatrix}, \quad
+    B = \begin{bmatrix}
+        b_{11} & b_{12} & \dots  & b_{1p} \\
+        b_{21} & b_{22} & \dots  & b_{2p} \\
+        \vdots & \vdots & \ddots & \vdots \\
+        b_{n1} & b_{n2} & \dots  & b_{np}
+    \end{bmatrix}
+$$
+
+The matrix multiplication of $A$ and $B$ is denoted as $C = A \cdot B$. The resulting matrix $C$ will have dimensions $m \times p$.
+
+    The entry $c_{ij}$ of the resulting matrix $C$ is computed as follows:
+
+    $$
+        c_{ij} = a_{i1} \cdot b_{1j} + a_{i2} \cdot b_{2j} + \dots + a_{in} \cdot b_{nj} = \sum_{k=1}^{n} a_{ik} \cdot b_{kj}
+    $$
+
+    The resulting matrix $C$ can be expressed as:
+
+    $$
+        C = \begin{bmatrix}
+            c_{11} & c_{12} & \dots  & c_{1p} \\
+            c_{21} & c_{22} & \dots  & c_{2p} \\
+            \vdots & \vdots & \ddots & \vdots \\
+            c_{m1} & c_{m2} & \dots  & c_{mp}
+        \end{bmatrix}
+    $$
+
+The definition above can be described as the following code:
+\begin{lstlisting}[language=Python, caption={Dot function}]
+@staticmethod
+def dot(A, B):
+    if A.is_matrix() and B.is_matrix():
+        a_rows, a_cols = A.shape()
+        b_rows, b_cols = B.shape()
+
+        if a_cols == b_rows:
+            result = [[0] * b_cols for _ in range(A.shape()[0])]
+            for i in range(A.shape()[0]):
+                for j in range(b_cols):
+                    for k in range(a_cols):
+                        result[i][j] += A.data[i][k] * B.data[k][j]
+            return Matrix(result)
+        else:
+            raise ValueError(
+                "Matrix dimensions do not match for dot product")
+    else:
+        raise ValueError("Dot product requires a Matrix object")
+\end{lstlisting}
+\vfill
+To parallelize this function, we simply delegate the calculation made in line 12 to an auxiliary function \textit{dot\_calc} defined in the \textit{tasks.py} file:
+\begin{lstlisting}[language=Python, caption={dot\_calc}]
+@ray.remote
+def dot_calc(A, B, i, j, k):
+    return A.data[i][k] * B.data[k][j]
+\end{lstlisting}
+
+\begin{lstlisting}[language=Python, caption={Parallelized dot function}]
+@staticmethod
+def dot(A, B):
+    if A.is_matrix() and B.is_matrix():
+        a_rows, a_cols = A.shape()
+        b_rows, b_cols = B.shape()
+
+        if a_cols == b_rows:
+            result = [[0] * b_cols for _ in range(a_rows)]
+            futures = []
+
+            for i in range(a_rows):
+                for j in range(b_cols):
+                    for k in range(a_cols):
+                        # result[i][j] += A.data[i][k] * B.data[k][j]
+                        futures.append(((i, j), t.dot_calc.remote(A, B, i, j, k)))
+
+            for future in futures:
+                i, j = future[0]
+                result[i][j] += ray.get(future[1])
+
+            return Matrix(result)
+        else:
+            raise ValueError(
+                "Matrix dimensions do not match for dot product")
+    else:
+        raise ValueError("Dot product requires a Matrix object")
+\end{lstlisting}
+
+\subsubsection{Determinant}
+\subsubsection{Rank}
+\subsubsection{Inverse}
+\pagebreak
 \section{Result of tests}

### 2023-09-19 latex/paper/main.tex
rank paper

@@
 \subsubsection{Determinant}
+
+\pagebreak
 \subsubsection{Rank}
-\subsubsection{Inverse}
+The \textbf{rank} of a matrix $A$, denoted as $\text{rank}(A)$, is defined as the maximum number of linearly independent rows or columns in the matrix.
+We chose to determine the rank of a matrix by using the minor criterion, also known as the \textbf{criterion of minors}.
+According to this criterion, the rank of a matrix is equal to the largest order of a non-zero determinant of any square submatrix within the given matrix.
+\\
+\\
+Let $A$ be a matrix of size $m \times n$, and let $j$ be the order of the largest non-zero determinant among all square submatrices of $A$.
+Then the rank of $A$, denoted as $\text{rank}(A)$, is equal to $j$.
+
+\begin{itemize}
+    \item{To apply the minor criterion, we calculate the determinants of all possible square submatrices of $A$, ranging from $1 \times 1$ to $\min(m,n) \times \min(m,n)$.
+The order of the largest non-zero determinant among these submatrices gives us the rank of $A$.}
+    \item{If there is a non-zero determinant of order $j$, but all determinants of order $k+1$ or higher are zero, then the rank of $A$ is $j$.}
+\end{itemize}
+This is then brought to code as the following:
+
+\begin{lstlisting}[language=Python, caption={Rank function}]
+def rank(self):
+    rows, cols = self.shape()
+    j1 = min(rows, cols)
+
+    for i in range(j1, 1, -1):
+        if self.get_square_submatrices(i) != -1:
+            return i    
+\end{lstlisting}
+The function \textit{get\_square\_submatrices(i)} is an auxiliary function used to calculate all the possible square submatrices of a matrix of order $i$:
+\begin{lstlisting}[language=Python, caption={get\_square\_submatrices}]
+def get_square_submatrices(self, order):
+    data = self.get()
+    rows, cols = self.shape()
+    submatrices = []
+
+    for start_row in range(rows - order + 1):
+        for start_col in range(cols - order + 1):
+            submatrix = []
+
+            for row in range(order):
+                submatrix.append(
+                    data[start_row + row][start_col:start_col + order])
+
+            submatrices.append(Matrix(submatrix))
+
+    return submatrices
+\end{lstlisting}
+This is then parallelized by distributing the computation of \textit{get\_square\_submatrices},  creating a further \textbf{remote} auxiliary function, \textit{get\_submatrix\_task}:
+\begin{lstlisting}[language=Python, caption={Rank function}]
+def get_square_submatrices(self, order):
+    data = self.get()
+    rows, cols = self.shape()
+    futures = []
+
+    for start_row in range(rows - order + 1):
+        for start_col in range(cols - order + 1):            
+            futures.append(t.get_submatrix_task.remote(start_row, start_col, order, data))
+
+    return ray.get(futures)
+\end{lstlisting}
+
+\begin{lstlisting}[language=Python, caption={get\_submatrix\_task}]
+@ray.remote
+def get_submatrix_task(start_row, start_col, order, data):
+        from matrix import Matrix
+        submatrix = []
+
+        for row in range(order):
+            submatrix.append(
+                data[start_row + row][start_col:start_col + order])
+
+        return submatrix
+\end{lstlisting}
+
 \pagebreak

### 2023-09-19 latex/paper/main.tex
minor fixes

@@
 This is then parallelized by distributing the computation of \textit{get\_square\_submatrices},  creating a further \textbf{remote} auxiliary function, \textit{get\_submatrix\_task}:
-\begin{lstlisting}[language=Python, caption={Rank function}]
+\begin{lstlisting}[language=Python, caption={get\_square\_submatrices}]
 def get_square_submatrices(self, order):

### 2023-09-29 distmat/main.py
tests draft

@@ -2,10 +2,58 @@ import ray
 import logging
 from matrix import Matrix
 import numpy as np
+import time
+import csv
+
+def test_dot_range(l, u, runs):
+    measures = []
+    sub_runs = []
+    for n in range(l, u):
+        for i in range(runs):
+            a = Matrix.random_int(16, 16, -10**n, 10**n)
+            b = Matrix.random_float(16, 16, -10**n, 10**n)
+
+            t0 = time.time()
+            Matrix.dot(a,b)
+            t1 = time.time()
+
+            sub_runs.append(t1-t0)
+            print(t1-t0)
+
+        measures.append(t1-t0)
+
+    return measures
+    
+def test_dot_size(l, u, runs):
+    measures = []
+    sub_runs = []
+    for n in range(l, u):
+        for i in range(runs):
+            a = Matrix.random_int(n, n, -99, 99)
+            b = Matrix.random_float(n, n, -99, 99)
+
+            t0 = time.time()
+            Matrix.dot(a,b)
+            t1 = time.time()
+
+            sub_runs.append(t1-t0)
+            print(t1-t0)
+
+        measures.append(sub_runs)
+
+    return measures
+
 
 if __name__ == "__main__":
-    a = Matrix.random_int(8, 8, -99, 99)
-    b = Matrix.random_float(8, 8, -99, 99)
+    if ray.is_initialized:
+        ray.shutdown()
+    ray.init(include_dashboard=True)
+
+    test_results = test_dot_size(2, 8, 4)    
+
+    with open("output.csv", "w") as f:
+        writer = csv.writer(f)
+        writer.writerows(test_results)
 
     # DOT PRODUCT
     # print("\ndistmat:\n", Matrix.dot(a,b))
@@ -20,10 +68,6 @@ if __name__ == "__main__":
     # print("distmat: ", a.rank())
 
     # INVERSE
-    print("\ndistmat:\n", a.inv())
-    print("numpy: \n", np.linalg.inv(np.array(a.get())))
-
-    if ray.is_initialized:
-        ray.shutdown()
-    ray.init(include_dashboard=True)
+    # print("\ndistmat:\n", a.inv())
+    # print("numpy: \n", np.linalg.inv(np.array(a.get())))
     

### 2023-09-29 distmat/main.py
plots

@@ -1,7 +1,5 @@
 import ray
-import logging
 from matrix import Matrix
-import numpy as np
 import time
 import csv
 
@@ -26,8 +24,11 @@ def test_dot_range(l, u, runs):
     
 def test_dot_size(l, u, runs):
     measures = []
-    sub_runs = []
+    
     for n in range(l, u):
+        sub_run = []
+        sub_run.append(n) # matrix size
+
         for i in range(runs):
             a = Matrix.random_int(n, n, -99, 99)
             b = Matrix.random_float(n, n, -99, 99)
@@ -36,10 +37,10 @@ def test_dot_size(l, u, runs):
             Matrix.dot(a,b)
             t1 = time.time()
 
-            sub_runs.append(t1-t0)
+            sub_run.append(t1-t0)
             print(t1-t0)
 
-        measures.append(sub_runs)
+        measures.append(sub_run)
 
     return measures
 
@@ -49,9 +50,9 @@ if __name__ == "__main__":
         ray.shutdown()
     ray.init(include_dashboard=True)
 
-    test_results = test_dot_size(2, 8, 4)    
+    test_results = test_dot_size(2, 16, 4)    
 
-    with open("output.csv", "w") as f:
+    with open("test_results/test_dot_size_2_16_4.csv", "w") as f:
         writer = csv.writer(f)
         writer.writerows(test_results)
 

### 2023-09-29 distmat/plots.py

@@ -1,0 +1,25 @@
+import csv
+import matplotlib.pyplot as plt
+import numpy as np
+
+# read data from csv
+data = []
+with open("output.csv", "r") as f:
+    reader = csv.reader(f)
+    for row in reader:
+        data.append([float(val) for val in row])
+
+# extract data
+sizes = [run[0] for run in data]
+means = [np.mean(runs) for runs in data]
+std_devs = [np.std(runs) for runs in data]
+
+# plot data
+plt.errorbar(sizes, means, yerr=std_devs, fmt='o-', label='Execution Time')
+plt.xlabel('Matrix Size (n)')
+plt.ylabel('Execution Time (s)')
+plt.title('Execution Time vs. Matrix Size')
+plt.grid()
+plt.legend()
+plt.savefig('test_dot_size_2_16_4.png')
+plt.show()

### 2023-09-30 latex/paper/main.tex
determinant

@@
 \lstdefinestyle{mystyle}{
+    aboveskip=15pt, % Adjust the space above the listing
+    belowskip=15pt, % Adjust the space below the listing
     backgroundcolor=\color{backcolour},   
@@
 $$
-
-The matrix multiplication of $A$ and $B$ is denoted as $C = A \cdot B$. The resulting matrix $C$ will have dimensions $m \times p$.
-
+The matrix multiplication of $A$ and $B$ is denoted as $C = A \cdot B$.
+The resulting matrix $C$ will have dimensions $m \times p$.
 The entry $c_{ij}$ of the resulting matrix $C$ is computed as follows:
-
     $$
@@
     $$
-
 The resulting matrix $C$ can be expressed as:
-
     $$
@@
     $$
-
+    \\
 The definition above can be described as the following code:
@@
 
+\pagebreak
 \subsubsection{Determinant}
+The \textbf{determinant} is a scalar value that is a function of the entries of a square matrix.
+It characterizes some properties of the matrix and the linear map represented by the matrix.
+In particular, the determinant is nonzero if and only if the matrix is \textbf{invertible}.
+\\\\
+To compute the determinant of matrices with order greater than 2 we are going to use the \textbf{Laplace method}.
+Let $A$ be a square matrix of size $n \times n$.
+The Laplace expansion of the determinant along the $i$th row is given by:
+
+$$
+    \det(A) = a_{i1} C_{i1} + a_{i2} C_{i2} + \ldots + a_{in} C_{in}
+$$
+where $C_{ij}$ denotes the cofactor of the element $a_{ij}$. The \textbf{cofactor} $C_{ij}$ is calculated as follows:
+
+$$
+    C_{ij} = (-1)^{i+j} \cdot \det(M_{ij})
+$$
+where $\det(M_{ij})$ represents the determinant of the submatrix obtained by deleting the $i$th row and $j$th column from matrix $A$.
+\\\\
+Using the Laplace method, the determinant can be calculated recursively by expanding along any row or column until a $2 \times 2$ matrix is reached, for which the determinant can be directly computed.
+It provides an alternative approach for determining the determinant of a matrix and can be particularly useful for matrices of larger sizes.
+It's an essential operation since it's used in the other operations of this study.
+\\\\
+The Laplace method can be transposed to code in the following way:
+\begin{lstlisting}[language=Python, caption={Determinant function}]
+def det(self):
+    if self.is_square():
+        data = self.get()
+        _, cols = self.shape()
+
+        if cols == 1:
+            return data[0][0]
+
+        elif cols == 2:
+            return (data[0][0] * data[1][1]) - (data[0][1] * data[1][0])
+
+        else:
+            det_value = 0
+
+            for j in range(cols):
+                minor = self.minor(0, j)
+                det_value += ((-1) ** j) * data[0][j] * minor.det()
+
+            return det_value
+
+    else:
+        raise ValueError("Cannot compute determinant of a non-square matrix")
+\end{lstlisting}
+The function is then parallelized by defining a variant of the \textit{minor} function, \textit{dist\_minor}, which distributes the workload of computing all minors of the matrix.
+
+\begin{lstlisting}[language=Python, caption={dist\_minor}]
+@ray.remote
+def dist_minor(A, i, j):
+    '''
+    Extract a minor matrix by removing the ith row and jth column
+    '''
+
+    data = A.get()
+    minor_data = [row[:j] + row[j + 1:]
+                  for row_idx, row in enumerate(data) if row_idx != i]
+
+    return Matrix(minor_data)
+\end{lstlisting}
+
+\begin{lstlisting}[language=Python, caption={Parallelized determinant function}]
+def det(self):
+if self.is_square():
+    data = self.get()
+    _, cols = self.shape()
+
+    if cols == 1:
+        return data[0][0]
+
+    elif cols == 2:
+        return (data[0][0] * data[1][1]) - (data[0][1] * data[1][0])
+
+    else:
+        det_value = 0
+        minors_futures = 
+            [Matrix.dist_minor.remote(self, 0, j) for j in range(cols)]
+
+            minors = ray.get(minors_futures)
+
+        for minor, j in zip(minors, range(cols)):
+            # minor = self.minor(0, j)
+            det_value += ((-1) ** j) * data[0][j] * minor.det()
+
+        return det_value
+
+else:
+    raise ValueError(
+        "Cannot compute determinant of a non-square matrix")
+\end{lstlisting}
 

### 2023-09-30 latex/paper/main.tex
paper ray clusters

@@
 
-\pagebreak
 \subsection{The Ray library}
@@
 
-\subsection{Ray clusters}
+\subsubsection{Ray clusters}
+To run Ray applications on multiple nodes you must first deploy a \textbf{Ray cluster}.
+A Ray cluster is a set of worker nodes connected to a common Ray head node.
+These nodes can be physical machines, virtual machines, or containers running on a cloud infrastructure or a cluster manager like Kubernetes.
+Ray clusters can be fixed-size, or they may autoscale up and down according to the resources requested by applications running on the cluster.
+\\\\
+Each Ray cluster typically has one special node called the \textbf{head node}.
+This is where the main driver program is executed, which manages the cluster, schedules tasks, and communicates with the other nodes.
+The head node also hosts the Ray Dashboard, a web-based interface for monitoring and managing the cluster.
+\\\\
+The other nodes in the cluster are known as \textbf{workers}.
+These nodes are responsible for executing tasks and running jobs.
+Workers can be distributed across multiple machines and are managed by the head node.
+They execute user-defined Python functions and can be scaled up or down dynamically based on demand.
+\\\\
+A ray cluster is set up using a \textit{yaml} file, where all the details about the cluster configuration are defined.
+Once the file is ready, we simply launch \textit{ray up cluster.yaml} to start a ray instance on all worker nodes over ssh.
+\begin{lstlisting}[caption={cluster.yaml}]
+    cluster_name: distmat
+
+    provider:
+      type: local
+      head_ip: "192.168.128.129"
+      worker_ips:
+        [
+          "192.168.128.210",
+          "192.168.128.211",
+          "192.168.128.212",
+          "192.168.128.213",
+          "192.168.128.214",
+          "192.168.128.220",
+          "192.168.128.221",
+          "192.168.128.222",
+          "192.168.128.223",
+          "192.168.128.224",
+        ]
+
+    auth:
+      ssh_user: alo
+
+    upscaling_speed: 1.0
+    idle_timeout_minutes: 5
+    head_start_ray_commands:
+      - ray stop
+      - ulimit -c unlimited && ray start --head --port=6379 --autoscaling-config=~/ray_bootstrap_config.yaml
+    worker_start_ray_commands:
+      - ray stop
+      - ray start --address=$RAY_HEAD_IP:6379
+
+\end{lstlisting}
 
@@
 Before implementing the parallelized algorithms for Ray, we write the code for the serial execution in the \textbf{Matrix} class (most of the functions defined in the class are auxiliary therefore were excluded for practical purposes):
+
 \begin{lstlisting}[language=Python, label=Matrix class]

### 2023-10-02 latex/paper/main.tex
paper inverse theory

@@
 auth:
-      ssh_user: alo
+    ssh_user: <user>
 
@@
 \end{lstlisting}
+
+\subsubsection{Cluster architecture}
 
@@
 \pagebreak
+\subsubsection{Inverse}
+The \textbf{inverse of a matrix} is a fundamental concept in linear algebra.
+For a square matrix $A$, if an inverse exists, it is denoted as $A^{-1}$.
+A matrix is invertible (or non-singular) if and only if its determinant is non-zero.
+The concept of the inverse of a matrix is closely related to solving linear systems of equations.
+Consider a system of linear equations represented in matrix form as $Ax = b$:
+$$
+\begin{array}{@{}cccc@{}}
+    A & x &  & b \\
+    \begin{bmatrix}
+        a_{11} & a_{12} & \cdots & a_{1n} \\
+        a_{21} & a_{22} & \cdots & a_{2n} \\
+        \vdots & \vdots & \ddots & \vdots \\
+        a_{m1} & a_{m2} & \cdots & a_{mn}
+    \end{bmatrix}
+    &
+    \begin{bmatrix}
+        x_1 \\
+        x_2 \\
+        \vdots \\
+        x_n
+    \end{bmatrix}
+    & = &
+    \begin{bmatrix}
+        b_1 \\
+        b_2 \\
+        \vdots \\
+        b_m
+    \end{bmatrix}
+\end{array}
+$$
+where:
+\begin{itemize}
+    \item{$A$ is the coefficient matrix}
+    \item{$x$ is the vector of variables we want to solve for}
+    \item{$b$ is the vector of constants on the right-hand side}
+\end{itemize}
+The \textbf{solution} for $x$ can be found using the inverse of $A$, hence $x = A^{-1}b$. To find the inverse of a matrix, one common method is to use the formula:
+$$
+    A^{-1} = \frac{1}{\det(A)} \cdot \text{adj}(A)
+$$
+where $\det(A)$ denotes the determinant of matrix $A$ and $\text{adj}(A)$ represents the \textbf{adjugate} of matrix $A$.
+The adjugate of matrix $A$ is obtained by taking the transpose of the \textbf{matrix of cofactors} of $A$.
+The cofactor $C_{ij}$ is calculated as:
+$$
+    C_{ij} = (-1)^{i+j} \cdot \det(M_{ij})
+$$
+where $\det(M_{ij})$ represents the determinant of the submatrix obtained by deleting the $i$th row and $j$th column from matrix $A$. The inverse of matrix $A$ satisfies the following condition:
+$$
+    A \cdot A^{-1} = A^{-1} \cdot A = I
+$$
+
 \section{Result of tests}

### 2023-10-02 latex/paper/main.tex
paper inv section

@@
 \definecolor{codepurple}{rgb}{0.58,0,0.82}
-\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+% \definecolor{backcolour}{rgb}{0.95,0.95,0.95}
+\definecolor{backcolour}{rgb}{1,1,1}
 
 \lstdefinestyle{mystyle}{
-    aboveskip=15pt, % Adjust the space above the listing
-    belowskip=15pt, % Adjust the space below the listing
     backgroundcolor=\color{backcolour},   
@@
     basicstyle=\ttfamily\footnotesize,
+    aboveskip=15pt, % Adjust the space above the listing
+    belowskip=15pt, % Adjust the space below the listing
     breakatwhitespace=false,         
@@
 $$
+\pagebreak
+\\
+We write the algorithm in Python as the following:
+\begin{lstlisting}[language=Python, label={Inverse function}]
+def inv(self):
+    if not self.is_square():
+        raise Exception("Matrix must be square")
+
+    elif self.det() == 0:
+        raise Exception("Matrix is not invertible")
+
+    else:
+        rows, cols = self.shape()
+        data = self.get()
+
+        det = self.det()
+
+        # special case for 2x2 matrix:
+        if rows == cols == 2:
+            return [[data[1][1] / det, -1 * data[0][1] / det],
+                    [-1 * data[1][0] / det, data[0][0] / det]]
+
+        # find matrix of cofactors
+        cof_matrix = []
+
+        for row in range(rows):
+            cof_row = []
+
+            for column in range(cols):
+                minor = self.minor(row, column)
+
+                cof_row.append(((-1)**(row + column)) * minor.det())
+
+            cof_matrix.append(cof_row)
+
+        cof_matrix = Matrix(cof_matrix).transpose()
+
+        cof_rows, cof_cols = cof_matrix.shape()
+        cof_data = cof_matrix.get()
+
+        for row in range(cof_rows):
+            for column in range(cof_cols):
+                cof_data[row][column] = cof_data[row][column] / det
+
+        return cof_matrix
+\end{lstlisting}
+To parallelize this function, we make use of the previously parallelized function \textit{det}, and define the new auxiliary functions \textit{inv\_cof\_matrix} and \textit{inv\_calc}:
+
+\begin{lstlisting}[language=Python, caption={inv\_calc}]
+    @ray.remote
+    def inv_calc(row, cof_cols, cof_data, det):
+        for col in range(cof_cols):
+            return (row, col, (cof_data[row][col] / det))
+    \end{lstlisting}
+\pagebreak
+\begin{lstlisting}[language=Python, caption={inv\_cof\_matrix}]
+@ray.remote
+def inv_cof_matrix(A, row, cols):
+
+    cof_row = []
+
+    for column in range(cols):
+        minor = A.minor(row, column)
+        cof_row.append(((-1)**(row + column)) * minor.det())
+
+    return cof_row
+\end{lstlisting}
+\begin{lstlisting}[language=Python, caption={Parallelized inverse function}]
+def inv(self):
+    if not self.is_square():
+        raise Exception("Matrix must be square")
+
+    elif self.det() == 0:
+        raise Exception("Matrix is not invertible")
+
+    else:
+        rows, cols = self.shape()
+        data = self.get()
+        det = self.det()
+
+        # special case for 2x2 matrix:
+        if rows == cols == 2:
+            m = [[data[1][1] / det, -1 * data[0][1] / det],
+                    [-1 * data[1][0] / det, data[0][0] / det]]
+
+            return Matrix([[round(i, 8) for i in j] for j in m])
+
+        # find matrix of cofactors
+        cof_rows_futures = [t.inv_cof_matrix.remote(self, row, cols) 
+                                for row in range(rows)]
+
+        cof_matrix = ray.get(cof_rows_futures)
+        cof_matrix = Matrix(cof_matrix).transpose()
+
+        cof_rows, cof_cols = cof_matrix.shape()
+        cof_data = cof_matrix.get()
+
+        data_futures = [t.inv_calc.remote(row, cof_cols, cof_data, det) 
+                            for row in range(cof_rows)]
+
+        for data in ray.get(data_futures):
+            for row, col, value in [data]:
+                cof_data[row][col] = value
+
+        return Matrix([[round(i, 8) for i in j] for j in cof_matrix])
+\end{lstlisting}
 

### 2023-10-03 distmat/main.py
dot test

@@ -2,27 +2,11 @@ import ray
 from matrix import Matrix
 import time
 import csv
-
-def test_dot_range(l, u, runs):
-    measures = []
-    sub_runs = []
-    for n in range(l, u):
-        for i in range(runs):
-            a = Matrix.random_int(16, 16, -10**n, 10**n)
-            b = Matrix.random_float(16, 16, -10**n, 10**n)
-
-            t0 = time.time()
-            Matrix.dot(a,b)
-            t1 = time.time()
-
-            sub_runs.append(t1-t0)
-            print(t1-t0)
-
-        measures.append(t1-t0)
-
-    return measures
     
 def test_dot_size(l, u, runs):
+    '''Make tests for matrices size l...u'''
+    max_n = (u-l)*5
+    counter = 0
     measures = []
     
     for n in range(l, u):
@@ -30,7 +14,7 @@ def test_dot_size(l, u, runs):
         sub_run.append(n) # matrix size
 
         for i in range(runs):
-            a = Matrix.random_int(n, n, -99, 99)
+            a = Matrix.random_int(n, n, -10**16, 10**16)
             b = Matrix.random_float(n, n, -99, 99)
 
             t0 = time.time()
@@ -38,11 +22,41 @@ def test_dot_size(l, u, runs):
             t1 = time.time()
 
             sub_run.append(t1-t0)
-            print(t1-t0)
+
+            counter += 1
+            print(f"run # {counter}/{max_n}: {t1-t0}")
 
         measures.append(sub_run)
 
     return measures
+
+def test_det_size(l, u, runs):
+    '''Make tests for matrices size l...u'''
+    max_n = (u-l)*5
+    counter = 0
+    measures = []
+    
+    for n in range(l, u):
+        sub_run = []
+        sub_run.append(n) # matrix size
+
+        for i in range(runs):
+            a = Matrix.random_int(n, n, -10**16, 10**16)
+            b = Matrix.random_float(n, n, -99, 99)
+
+            t0 = time.time()
+            Matrix.dot(a,b)
+            t1 = time.time()
+
+            sub_run.append(t1-t0)
+
+            counter += 1
+            print(f"run # {counter}/{max_n}: {t1-t0}")
+
+        measures.append(sub_run)
+
+    return measures
+
 
 
 if __name__ == "__main__":
@@ -50,9 +64,10 @@ if __name__ == "__main__":
         ray.shutdown()
[…]

### 2023-10-03 distmat/plots.py

@@ -1,10 +1,11 @@
 import csv
 import matplotlib.pyplot as plt
 import numpy as np
+import sys
 
 # read data from csv
 data = []
-with open("output.csv", "r") as f:
+with open(sys.argv[1], "r") as f:
     reader = csv.reader(f)
     for row in reader:
         data.append([float(val) for val in row])
@@ -21,5 +22,5 @@ plt.ylabel('Execution Time (s)')
 plt.title('Execution Time vs. Matrix Size')
 plt.grid()
 plt.legend()
-plt.savefig('test_dot_size_2_16_4.png')
+# plt.savefig(f"{sys.argv[1]}.png")
 plt.show()

### 2023-10-05 distmat/tasks.py
updated inv in tasks

@@ -17,12 +17,14 @@ def get_submatrix_task(start_row, start_col, order, data):
 
 @ray.remote
 def inv_cof_matrix(A, row, cols):
-
+    from matrix import Matrix
+    
     cof_row = []
+    minor_futures = [Matrix.dist_minor.remote(A, row, col) for col in range(cols)]
+    minors = ray.get(minor_futures)
 
-    for column in range(cols):
-        minor = A.minor(row, column)
-        cof_row.append(((-1)**(row + column)) * minor.det())
+    for col,minor in zip(range(cols), minors):
+        cof_row.append(((-1)**(row + col)) * minor.det())
 
     return cof_row
 

### 2023-10-05 latex/paper/main.tex
comments
updated paper

@@
 \subsubsection{Cluster architecture}
+/scrivere informazioni riguardo il cluster, come le specifiche tecniche e la quantita di macchine/
 
@@
 def inv_cof_matrix(A, row, cols):
+    from matrix import Matrix
 
     cof_row = []
+    minor_futures = [Matrix.dist_minor.remote(A, row, col) for col in range(cols)]
+    minors = ray.get(minor_futures)
 
-    for column in range(cols):
-        minor = A.minor(row, column)
-        cof_row.append(((-1)**(row + column)) * minor.det())
+    for col,minor in zip(range(cols), minors):
+        cof_row.append(((-1)**(row + col)) * minor.det())
 

### 2023-10-05 distmat/main.py

@@ -3,9 +3,9 @@ from matrix import Matrix
 import time
 import csv
     
-def test_dot_size(l, u, runs):
+def test_dot(l, u, runs):
     '''Make tests for matrices size l...u'''
-    max_n = (u-l)*5
+    max_n = (u-l)*runs
     counter = 0
     measures = []
     
@@ -14,7 +14,7 @@ def test_dot_size(l, u, runs):
         sub_run.append(n) # matrix size
 
         for i in range(runs):
-            a = Matrix.random_int(n, n, -10**16, 10**16)
+            a = Matrix.random_int(n, n, -10**8, 10**8)
             b = Matrix.random_float(n, n, -99, 99)
 
             t0 = time.time()
@@ -30,9 +30,9 @@ def test_dot_size(l, u, runs):
 
     return measures
 
-def test_det_size(l, u, runs):
+def test_det(l, u, runs):
     '''Make tests for matrices size l...u'''
-    max_n = (u-l)*5
+    max_n = (u-l)*runs
     counter = 0
     measures = []
     
@@ -41,11 +41,10 @@ def test_det_size(l, u, runs):
         sub_run.append(n) # matrix size
 
         for i in range(runs):
-            a = Matrix.random_int(n, n, -10**16, 10**16)
-            b = Matrix.random_float(n, n, -99, 99)
+            a = Matrix.random_float(n, n, -10**8, 10**8)
 
             t0 = time.time()
-            Matrix.dot(a,b)
+            a.det()
             t1 = time.time()
 
             sub_run.append(t1-t0)
@@ -57,33 +56,70 @@ def test_det_size(l, u, runs):
 
     return measures
 
+def test_rank(l, u, runs):
+    '''Make tests for matrices size l...u'''
+    max_n = (u-l)*runs
+    counter = 0
+    measures = []
+    
+    for n in range(l, u):
+        sub_run = []
+        sub_run.append(n) # matrix size
 
+        for i in range(runs):
+            a = Matrix.random_float(n, n, -10**8, 10**8)
+
+            t0 = time.time()
+            a.rank()
+            t1 = time.time()
+
+            sub_run.append(t1-t0)
+
+            counter += 1
+            print(f"run # {counter}/{max_n}: {t1-t0}")
+
+        measures.append(sub_run)
+
+    return measures
+
+def test_inv(l, u, runs):
+    '''Make tests for matrices size l...u'''
[…]

### 2023-10-05 distmat/main.py
test inv

@@ -113,13 +113,13 @@ if __name__ == "__main__":
         ray.shutdown()
     ray.init(include_dashboard=True)
 
-    # test_results = test_inv(2, 9, 3)
+    test_results = test_inv(2, 8, 3)
 
-    # with open("test_results/test_inv_2_20_3.csv", "w") as f:
-    #     writer = csv.writer(f)
-    #     writer.writerows(test_results)
+    with open("test_results/test_inv_2_8_3.csv", "w") as f:
+        writer = csv.writer(f)
+        writer.writerows(test_results)
 
-    a = Matrix.random_float(8, 8, -10**8, 10**8)
-    t0 = time.time()
-    a.inv()
-    print(time.time() - t0)
+    # a = Matrix.random_float(8, 8, -10**8, 10**8)
+    # t0 = time.time()
+    # a.inv()
+    # print(time.time() - t0)

### 2023-10-07 distmat/main.py
serial tests and plotting

@@ -1,14 +1,16 @@
 import ray
-from matrix import Matrix
+from matrix_serial import Matrix
 import time
 import csv
     
-def test_dot(l, u, runs):
+def test_dot(l, u, runs, filename):
     '''Make tests for matrices size l...u'''
     max_n = (u-l)*runs
     counter = 0
     measures = []
     
+    print(f"running test_dot with l={l}, u={u}, runs={runs}")
+
     for n in range(l, u):
         sub_run = []
         sub_run.append(n) # matrix size
@@ -28,14 +30,20 @@ def test_dot(l, u, runs):
 
         measures.append(sub_run)
 
+    with open(f"{filename}_{l}_{u}_{runs}.csv", "w") as f:
+        writer = csv.writer(f)
+        writer.writerows(measures)
+
     return measures
 
-def test_det(l, u, runs):
+def test_det(l, u, runs, filename):
     '''Make tests for matrices size l...u'''
     max_n = (u-l)*runs
     counter = 0
     measures = []
     
+    print(f"running test_det with l={l}, u={u}, runs={runs}")
+
     for n in range(l, u):
         sub_run = []
         sub_run.append(n) # matrix size
@@ -53,15 +61,21 @@ def test_det(l, u, runs):
             print(f"run # {counter}/{max_n}: {t1-t0}")
 
         measures.append(sub_run)
+    
+    with open(f"{filename}_{l}_{u}_{runs}.csv", "w") as f:
+        writer = csv.writer(f)
+        writer.writerows(measures)
 
     return measures
 
-def test_rank(l, u, runs):
+def test_rank(l, u, runs, filename):
     '''Make tests for matrices size l...u'''
     max_n = (u-l)*runs
     counter = 0
     measures = []
     
+    print(f"running test_rank with l={l}, u={u}, runs={runs}")
+
     for n in range(l, u):
         sub_run = []
         sub_run.append(n) # matrix size
@@ -80,14 +94,20 @@ def test_rank(l, u, runs):
 
         measures.append(sub_run)
 
+    with open(f"{filename}_{l}_{u}_{runs}.csv", "w") as f:
+        writer = csv.writer(f)
+        writer.writerows(measures)
+
     return measures
 
-def test_inv(l, u, runs):
+def test_inv(l, u, runs, filename):
     '''Make tests for matrices size l...u'''
     max_n = (u-l)*runs
     counter = 0
     measures = []
     
[…]

### 2023-10-07 distmat/matrix_serial.py

@@ -1,0 +1,262 @@
+import random
+
+class Matrix:
+    def __init__(self, data):
+        if isinstance(data, list):
+            self.data = data
+
+        else:
+            raise ValueError("Input must be a list")
+
+    def __str__(self):
+        tmp = ""
+        for row in self.data:
+            tmp = tmp + str(row) + '\n'
+        return tmp
+
+    def __getitem__(self, index):
+        return self.data[index]
+
+    def __setitem__(self, index, value):
+        self.data[index] = value
+
+    def shape(self):
+        num_rows = len(self.data)
+        num_cols = len(self.data[0]) if num_rows > 0 else 0
+        return (num_rows, num_cols)
+
+    def get(self):
+        return self.data
+
+    @staticmethod
+    def random_int(rows, cols, l, u):
+        if rows > 1 and cols > 1:
+            data = [[random.randint(l, u) for _ in range(rows)]
+                    for _ in range(cols)]
+
+        elif rows == 1 and cols > 1:
+            data = [random.randint(l, u) for _ in range(cols)]
+
+        elif rows > 1 and cols == 1:
+            data = [[random.randint(l, u)]
+                    for _ in range(cols) for _ in range(rows)]
+
+        return Matrix(data)
+
+    @staticmethod
+    def random_float(rows, cols, l, u):
+        if rows > 1 and cols > 1:
+            data = [[random.uniform(l, u) for _ in range(rows)]
+                    for _ in range(cols)]
+
+        elif rows == 1 and cols > 1:
+            data = [random.uniform(l, u) for _ in range(cols)]
+
+        elif rows > 1 and cols == 1:
+            data = [[random.uniform(l, u)]
+                    for _ in range(cols) for _ in range(rows)]
+
+        return Matrix(data)
+
+    def is_matrix(self):
+        rows, cols = self.shape()
+        return rows > 1 and cols > 1
+
+    def is_square(self):
+        rows, cols = self.shape()
+        return rows == cols
+
+    def is_vector(self):
+        rows, cols = self.shape()
+        return rows == 1 ^ cols == 1
+
+    def is_row_vector(self):
+        rows, cols = self.shape()
+        return rows == 1 and cols > 1
+
+    def is_col_vector(self):
+        rows, cols = self.shape()
+        return cols == 1 and rows > 1
+
+    def is_scalar(self):
+        rows, cols = self.shape()
+        return rows == cols and cols == 1
+
+    def transpose(self):
+        data = self.get()
[…]

### 2023-10-07 distmat/plots.py

@@ -1,26 +1,92 @@
 import csv
 import matplotlib.pyplot as plt
 import numpy as np
-import sys
 
-# read data from csv
-data = []
-with open(sys.argv[1], "r") as f:
-    reader = csv.reader(f)
-    for row in reader:
-        data.append([float(val) for val in row])
+def serial_parallel_comparison(csv_name):
+    # Read data from serial CSV file
+    serial_data = []
+    with open(f"test_results/serial/{csv_name}", "r") as f:
+        reader = csv.reader(f)
+        for row in reader:
+            serial_data.append([float(val) for val in row])
 
-# extract data
-sizes = [run[0] for run in data]
-means = [np.mean(runs) for runs in data]
-std_devs = [np.std(runs) for runs in data]
+    # Read data from parallel CSV file
+    parallel_data = []
+    with open(f"test_results/parallel/{csv_name}", "r") as f:
+        reader = csv.reader(f)
+        for row in reader:
+            parallel_data.append([float(val) for val in row])
 
-# plot data
-plt.errorbar(sizes, means, yerr=std_devs, fmt='o-', label='Execution Time')
-plt.xlabel('Matrix Size (n)')
-plt.ylabel('Execution Time (s)')
-plt.title('Execution Time vs. Matrix Size')
-plt.grid()
-plt.legend()
-# plt.savefig(f"{sys.argv[1]}.png")
-plt.show()
+    # Extract serial data
+    serial_sizes = [run[0] for run in serial_data]
+    serial_means = [np.mean(runs) for runs in serial_data]
+    serial_std_devs = [np.std(runs) for runs in serial_data]
+
+    # Extract parallel data
+    parallel_sizes = [run[0] for run in parallel_data]
+    parallel_means = [np.mean(runs) for runs in parallel_data]
+    parallel_std_devs = [np.std(runs) for runs in parallel_data]
+
+    # Plot data
+    plt.errorbar(serial_sizes, serial_means, yerr=serial_std_devs, fmt='o-', label='Serial Execution Time')
+    plt.errorbar(parallel_sizes, parallel_means, yerr=parallel_std_devs, fmt='x-', label='Parallel Execution Time')
+    plt.xlabel('Matrix Size (n)')
+    plt.ylabel('Execution Time (s)')
+    plt.title('Serial vs. Parallel Execution Time')
+    plt.grid()
+    plt.legend()
+    plt.show()
+
+def rank_serial_parallel_comparison(csv_name):
[…]

### 2023-10-07 distmat/plots.py
figures

@@ -2,7 +2,7 @@ import csv
 import matplotlib.pyplot as plt
 import numpy as np
 
-def serial_parallel_comparison(csv_name):
+def serial_parallel_comparison(csv_name, func_name):
     # Read data from serial CSV file
     serial_data = []
     with open(f"test_results/serial/{csv_name}", "r") as f:
@@ -32,9 +32,10 @@ def serial_parallel_comparison(csv_name):
     plt.errorbar(parallel_sizes, parallel_means, yerr=parallel_std_devs, fmt='x-', label='Parallel Execution Time')
     plt.xlabel('Matrix Size (n)')
     plt.ylabel('Execution Time (s)')
-    plt.title('Serial vs. Parallel Execution Time')
+    plt.title(f'Serial vs. parallel execution time ({func_name})')
     plt.grid()
     plt.legend()
+    plt.savefig(csv_name.replace("csv", "png"))
     plt.show()
 
 def rank_serial_parallel_comparison(csv_name):
@@ -77,15 +78,17 @@ def rank_serial_parallel_comparison(csv_name):
 
     plt.xlabel('Matrix Size (n)')
     plt.ylabel('Execution Time (s)')
-    plt.title('Serial vs. Parallel Execution Time')
+    plt.title('Serial vs. parallel execution time (rank)')
     plt.grid()
     plt.legend()
+    plt.savefig(csv_name.replace("csv", "png"))
     plt.show()
 
+
 # Call the function with your CSV file name
-serial_parallel_comparison("det_2_10_3.csv")
-serial_parallel_comparison("dot_2_30_5.csv")
-serial_parallel_comparison("inv_2_8_3.csv")
+serial_parallel_comparison("det_2_10_3.csv", "determinant")
+serial_parallel_comparison("dot_2_30_5.csv", "dot product")
+serial_parallel_comparison("inv_2_8_3.csv", "inverse")
 
 
 rank_serial_parallel_comparison("rank_2_500_5.csv")

### 2023-10-07 latex/paper/main.tex
introduction, case study section

@@
 
-\pagebreak
-\section{Case study}
-
-\section{Implementation}
 \subsection{Python}
@@
 It is widely used for a variety of applications, including web development, data analysis, artificial intelligence, scientific computing, and automation.
-It supports multiple programming paradigms, including procedural, object-oriented, and functional programming, and its large and active community contributes to a vast ecosystem of libraries and frameworks, making it a powerful tool for a wide range of tasks.
-It is one of the two languages supported by Ray, alongside Java.
+Python's strength lies in its support for multiple programming paradigms, including procedural, object-oriented, and functional programming.
+Its large and active community contributes to a vast ecosystem of libraries and frameworks, making it a powerful tool for a wide range of tasks.
+\\\\
+In the context of this paper, Python serves as the primary programming language for conducting tests and implementing linear algebra operations.
+Additionally, Matplotlib is utilized for graphing and visualizing the results of these tests.
+\\\\
+The object-oriented programming (OOP) paradigm has been employed in this research to represent matrices.
+This approach enhances code organization and readability while facilitating the implementation of complex linear algebra operations.
+\\\\
+It's worth noting that Python is one of the two languages supported by Ray, with the other being Java, making it a suitable choice for distributed computing and parallelization.
 
+\section{Case study}
+
+In this section, we present the case study conducted to evaluate the performance of distributed computation of linear algebra operations using Ray 2.4.0 on a cluster comprised of 10 computers with specifications provided by the University of Messina.
+The experiments include the caluclation of determinant, dot product, matrix inversion, and rank.
+For both serial and parallel execution, we conducted tests with increasing matrix sizes and repetitions to gather comprehensive performance data.
+\\\\
+The cluster used for our experiments is composed of ten computers with the following specifications, provided by the University of Messina:
+\begin{itemize}
+    \item {RAM: 8GB}
+    \item {CPU: DA INSERIRE}
+\end{itemize}
+\pagebreak
+\subsection{Execution of tests}
+For our experiments, we considered the following parameters:
+\begin{itemize}
+    \item \textbf{Determinant calculation:} Matrices of sizes $n = 2 \dots 10$ were used, with each $n$ value being executed three times. The average execution time was recorded for analysis.
+    \item \textbf{Dot product:} Matrices of sizes $n = 2 \dots 30$ were employed, with each $n$ value being executed five times.
+    \item \textbf{Inverse:} Matrices of sizes $n = 2 \dots 8$ were utilized, with each $n$ value being executed three times.
+    \item \textbf{Rank:} Matrices of sizes $n = 2 \dots 500$ were tested, with each $n$ value being executed five times.
+\end{itemize}
+\vfill
+For each type of operation, the code iterates through different matrix sizes, executing the operation multiple times for each size to capture variations in performance, and recording the execution times for analysis.
+The results of these tests are stored in separate CSV files, organized by operation type and parameter configuration.
+This is achieved with the following Python script:
+\pagebreak
+\begin{lstlisting}[language=Python, label=main.py]
+import ray
+from matrix_serial import Matrix
+import time
+import csv
+
+def test_dot(l, u, runs, filename):
+    '''Make tests for matrices size l...u'''
+    max_n = (u-l)*runs
+    counter = 0
+    measures = []
+
+    print(f"running test_dot with l={l}, u={u}, runs={runs}")
+
+    for n in range(l, u):
+        sub_run = []
+        sub_run.append(n) # matrix size
+
+        for i in range(runs):
+            a = Matrix.random_int(n, n, -10**8, 10**8)
+            b = Matrix.random_float(n, n, -99, 99)
+
+            t0 = time.time()
+            Matrix.dot(a,b)
+            t1 = time.time()
+
+            sub_run.append(t1-t0)
+
+            counter += 1
+            print(f"run # {counter}/{max_n}: {t1-t0}")
+
+        measures.append(sub_run)
+
+    with open(f"{filename}_{l}_{u}_{runs}.csv", "w") as f:
+        writer = csv.writer(f)
+        writer.writerows(measures)
+
+    return measures
+
+def test_det(l, u, runs, filename):
+    [...]
+def test_rank(l, u, runs, filename):
+    [...]
+def test_inv(l, u, runs, filename):
+    [...]
+
+if __name__ == "__main__":
+    if ray.is_initialized:
+        ray.shutdown()
+    ray.init(include_dashboard=True)
+
+    test_det(2, 10, 3, "test_results/[[serial/parallel]]/det")
+    test_dot(2, 30, 5, "test_results/[[serial/parallel]]/dot")
+    test_inv(2, 8, 3, "test_results/[[serial/parallel]]/inv")
+    test_rank(2, 500, 5, "test_results/[[serial/parallel]]/rank")
+    \end{lstlisting}
+
+\subsection{Graphing the results}
+The code loads CSV files onto defined variables and then generates plots with error bars to visualize the performance differences between serial and parallel execution using \textit{matplotlib}.
+The primary goal is to assess the efficiency and scalability of these operations when executed in parallel.
+The resulting plots are saved as PNG files for analysis.
+\\\\
+A different function was defined for the plotting of rank, downsampling the data. This is because of its particular case of increasing standard deviation value, thus legibility reasons.
+
+\begin{lstlisting}[language=Python, label=plots.py]
+import csv
+import matplotlib.pyplot as plt
+import numpy as np
+
+def serial_parallel_comparison(csv_name, func_name):
+    # Read data from serial CSV file
+    serial_data = []
+    with open(f"test_results/serial/{csv_name}", "r") as f:
+        reader = csv.reader(f)
+        for row in reader:
+            serial_data.append([float(val) for val in row])
+
+    # Read data from parallel CSV file
+    parallel_data = []
+    with open(f"test_results/parallel/{csv_name}", "r") as f:
+        reader = csv.reader(f)
+        for row in reader:
+            parallel_data.append([float(val) for val in row])
+
+    # Extract serial data
+    serial_sizes = [run[0] for run in serial_data]
+    serial_means = [np.mean(runs) for runs in serial_data]
+    serial_std_devs = [np.std(runs) for runs in serial_data]
+
+    # Extract parallel data
+    parallel_sizes = [run[0] for run in parallel_data]
+    parallel_means = [np.mean(runs) for runs in parallel_data]
+    parallel_std_devs = [np.std(runs) for runs in parallel_data]
+
+    # Plot data
+    plt.errorbar(serial_sizes, serial_means, yerr=serial_std_devs, fmt='o-', label='Serial Execution Time')
+    plt.errorbar(parallel_sizes, parallel_means, yerr=parallel_std_devs, fmt='x-', label='Parallel Execution Time')
+    plt.xlabel('Matrix Size (n)')
+    plt.ylabel('Execution Time (s)')
+    plt.title(f'Serial vs. parallel execution time ({func_name})')
+    plt.grid()
+    plt.legend()
+    plt.savefig(csv_name.replace("csv", "png"))
+    plt.show()
+
+def rank_serial_parallel_comparison(csv_name):
+    [...]
+    # Downsample the data
+    n = 10
+    serial_sizes = serial_sizes[::n]
+    serial_means = serial_means[::n]
+    serial_std_devs = serial_std_devs[::n]
+    parallel_sizes = parallel_sizes[::n]
+    parallel_means = parallel_means[::n]
+    parallel_std_devs = parallel_std_devs[::n]
+    [...]
+
+serial_parallel_comparison("det_2_10_3.csv", "determinant")
+serial_parallel_comparison("dot_2_30_5.csv", "dot product")
+serial_parallel_comparison("inv_2_8_3.csv", "inverse")
+rank_serial_parallel_comparison("rank_2_500_5.csv")
+\end{lstlisting}
+\section{Implementation}
 \subsection{Ray}
@@
 cluster_name: distmat
-
 provider:
@@
     ]
-
 auth:
     ssh_user: <user>
-
 upscaling_speed: 1.0
@@
     - ray start --address=$RAY_HEAD_IP:6379
-
 \end{lstlisting}
-
-\subsubsection{Cluster architecture}
-/scrivere informazioni riguardo il cluster, come le specifiche tecniche e la quantita di macchine/
-
 \pagebreak
@@
 
-\section{Result of tests}
-
 \section{Conclusion}

### 2023-10-09 latex/paper/main.tex
detailed graphs and det results

@@
 % This title page is capable of being compiled as is. This is not useful for 
-% including it in another document. To do this, you have two options: 
+% including it in another document. To do this, we have two options: 
 %
 % 1) Copy/paste everything between \begin{document} and \end{document} 
-% starting at \begin{titlepage} and paste this into another LaTeX file where you 
+% starting at \begin{titlepage} and paste this into another LaTeX file where we 
 % want your title page.
@@
 % 2) Remove everything outside the \begin{titlepage} and \end{titlepage} and 
-% move this file to the same directory as the LaTeX file you wish to add it to. 
-% Then add \input{./title_page_1.tex} to your LaTeX file where you want your
+% move this file to the same directory as the LaTeX file we wish to add it to. 
+% Then add \input{./title_page_1.tex} to your LaTeX file where we want your
 % title page.
@@
 
-    % If you don't want a supervisor, uncomment the two lines below and remove the section above
+    % If we don't want a supervisor, uncomment the two lines below and remove the section above
     %\Large \emph{Author:}\\
-    %John \textsc{Smith}\\[3cm] % Your name
+    %John \textsc{Smith}\\[3cm] % your name
 
@@
 
-    {\large \today}\\[2cm] % Date, change the \today to a set date if you want to be precise
+    {\large \today}\\[2cm] % Date, change the \today to a set date if we want to be precise
 
@@
 \begin{itemize}
-    \item \textbf{Faster execution}: With Ray you can use multiple machines and CPUs to compute operations.
+    \item \textbf{Faster execution}: With Ray we can use multiple machines and CPUs to compute operations.
 This significantly reduces the computation time compared to a single machine.
@@
     \item \textbf{Scalability}: As the size of the matrices grows, a single machine may struggle to handle the computational demands due to memory limitations or processing power constraints.
-Ray allows you to scale horizontally by adding more machines to the distributed setup.
+Ray allows we to scale horizontally by adding more machines to the distributed setup.
     \item \textbf{Fault tolerance}: Ray offers fault tolerance mechanisms that ensure the continuity of computation even in the presence of failures.
@@
     \item \textbf{Resource utilization}: With Ray, each machine contributes its processing power and memory capacity to the overall computation.
-This efficient utilization of resources allows you to make the most of the available hardware infrastructure, compared to a single machine that may be underutilized.
+This efficient utilization of resources allows we to make the most of the available hardware infrastructure, compared to a single machine that may be underutilized.
 \end{itemize}
@@
 To parallelize a function with Ray, import Ray and and initialize it with \textit{ray.init()}.
-Then decorate the function with \textit{@ray.remote} to declare that you want to run this function remotely.
+Then decorate the function with \textit{@ray.remote} to declare that we want to run this function remotely.
 Lastly, call the function with \textit{.remote()} instead of calling it normally.
-This remote call yields a future, a Ray object reference, that you can then fetch with \textit{ray.get}:
+This remote call yields a future, a Ray object reference, that we can then fetch with \textit{ray.get}:
 
@@
 \subsubsection{Ray clusters}
-To run Ray applications on multiple nodes you must first deploy a \textbf{Ray cluster}.
+To run Ray applications on multiple nodes we must first deploy a \textbf{Ray cluster}.
 A Ray cluster is a set of worker nodes connected to a common Ray head node.
@@
 \end{lstlisting}
+\pagebreak
+
+\section{Results of tests}
+
+\subsection{Determinant}
+The \textbf{serial} execution times are generally faster than the parallel execution times for smaller matrix sizes (2 and 3), with serial execution times measured in microseconds.
+However, as the matrix size increases, serial execution times also increase due to higher complexity of determinant calculations for larger matrices.
+Notably, the standard deviation for serial execution times also rises with matrix size, indicating increased variability in execution times for larger matrices in serial execution.
+\\\\
+Conversely, for smaller matrix sizes (2 and 3), \textbf{parallel} execution outperforms serial execution, with execution times measured in microseconds ( $10^{-6}$ seconds).
+Nevertheless, as the matrix size grows (4 and beyond), parallel execution times increase significantly, suggesting that the overhead associated with parallelization may begin to outweigh the benefits of parallelism for larger matrices.
+Importantly, the standard deviation for parallel execution times remains relatively stable as matrix size increases, indicating consistent results across multiple runs.
+
+\begin{figure}
+    \includegraphics[width=0.8\textwidth, keepaspectratio, center]{img/det_2_10_3.png}
+    \label{fig:det}
+    \caption{Determinant}
+
+    \includegraphics[width=0.8\textwidth, keepaspectratio, center]{img/det_2_10_3_detail.png}
+    \label{fig:det_detail}
+    \caption{Detailed plot}
+\end{figure}
+
+\subsection{Dot product}
+
+\begin{figure}
+    \includegraphics[width=0.8\textwidth, keepaspectratio, center]{img/dot_2_30_5.png}
+    \label{fig:dot}
+    \caption{Dot product}
+
+    \includegraphics[width=0.8\textwidth, keepaspectratio, center]{img/dot_2_30_5_detail.png}
+    \label{fig:dot_detail}
+    \caption{Detailed plot}
+\end{figure}
+
+\subsection{Inverse}
+
+\begin{figure}
+    \includegraphics[width=0.8\textwidth, keepaspectratio, center]{img/inv_2_8_3.png}
+    \label{fig:inv}
+    \caption{Inverse}
+
+    \includegraphics[width=0.8\textwidth, keepaspectratio, center]{img/inv_2_8_3_detail.png}
+    \label{fig:inv_detail}
+    \caption{Detailed plot}
+\end{figure}
+
+\subsection{Rank}
+\begin{figure}
+    \includegraphics[width=0.8\textwidth, keepaspectratio, center]{img/rank_2_500_5.png}
+    \label{fig:rank}
+    \caption{Rank}
+\end{figure}
 

### 2023-10-09 latex/paper/main.tex
figure placement

@@
 
+\usepackage{float}
 \usepackage[margin=3cm]{geometry}
@@
 
-\begin{figure}
-    \includegraphics[width=0.8\textwidth, keepaspectratio, center]{img/det_2_10_3.png}
+\begin{figure}[H]
+    \includegraphics[width=0.8\textwidth, keepaspectratio]{img/det_2_10_3.png}
+    \centering
     \label{fig:det}
@@
 
-    \includegraphics[width=0.8\textwidth, keepaspectratio, center]{img/det_2_10_3_detail.png}
+    \includegraphics[width=0.8\textwidth, keepaspectratio]{img/det_2_10_3_detail.png}
+    \centering
     \label{fig:det_detail}
@@
 
-\begin{figure}
-    \includegraphics[width=0.8\textwidth, keepaspectratio, center]{img/dot_2_30_5.png}
+\begin{figure}[H]
+    \includegraphics[width=0.8\textwidth, keepaspectratio]{img/dot_2_30_5.png}
+    \centering
     \label{fig:dot}
@@
 
-    \includegraphics[width=0.8\textwidth, keepaspectratio, center]{img/dot_2_30_5_detail.png}
+    \includegraphics[width=0.8\textwidth, keepaspectratio]{img/dot_2_30_5_detail.png}
+    \centering
     \label{fig:dot_detail}
@@
 
-\begin{figure}
-    \includegraphics[width=0.8\textwidth, keepaspectratio, center]{img/inv_2_8_3.png}
+\begin{figure}[H]
+    \includegraphics[width=0.8\textwidth, keepaspectratio]{img/inv_2_8_3.png}
+    \centering
     \label{fig:inv}
@@
 
-    \includegraphics[width=0.8\textwidth, keepaspectratio, center]{img/inv_2_8_3_detail.png}
+    \includegraphics[width=0.8\textwidth, keepaspectratio]{img/inv_2_8_3_detail.png}
+    \centering
     \label{fig:inv_detail}
@@
 \subsection{Rank}
-\begin{figure}
-    \includegraphics[width=0.8\textwidth, keepaspectratio, center]{img/rank_2_500_5.png}
+\begin{figure}[H]
+    \includegraphics[width=0.8\textwidth, keepaspectratio]{img/rank_2_500_5.png}
+    \centering
     \label{fig:rank}

### 2023-10-09 latex/paper/main.tex
dot product analysis

@@
 \subsection{Determinant}
-The \textbf{serial} execution times are generally faster than the parallel execution times for smaller matrix sizes (2 and 3), with serial execution times measured in microseconds.
-However, as the matrix size increases, serial execution times also increase due to higher complexity of determinant calculations for larger matrices.
-Notably, the standard deviation for serial execution times also rises with matrix size, indicating increased variability in execution times for larger matrices in serial execution.
+The \textbf{serial} execution times are generally faster than the \textbf{parallel} execution times, with both serial and parallel measured in microseconds ($10^{-6}$ seconds) for smaller matrices ($n \le 3$).
+As the matrix size increases, serial execution times also increase due to the higher complexity of determinant calculations.
+This trend is consistent with the observation that \textbf{parallel execution is generally slower than serial}, and this performance gap tends to widen as the matrix sizes increase.
+This suggests that the overhead associated with parallelization may begin to outweigh the benefits of parallelism for larger matrices.
 \\\\
-Conversely, for smaller matrix sizes (2 and 3), \textbf{parallel} execution outperforms serial execution, with execution times measured in microseconds ( $10^{-6}$ seconds).
-Nevertheless, as the matrix size grows (4 and beyond), parallel execution times increase significantly, suggesting that the overhead associated with parallelization may begin to outweigh the benefits of parallelism for larger matrices.
-Importantly, the standard deviation for parallel execution times remains relatively stable as matrix size increases, indicating consistent results across multiple runs.
+The standard deviation for serial execution times also rises with matrix size, indicating increased variability in execution times for larger matrices in serial execution, while the standard deviation for parallel execution times remains relatively stable as matrix size increases, indicating consistent results across multiple runs.
 
@@
 \subsection{Dot product}
-
+The parallel execution times for the dot product function exhibit a similar trend compared to the determinant function.
+In this case, the serial execution is generally faster and increases linearly as the matrix size grows.
+The execution times for both serial and parallel executions are measured in microseconds ($10^{-6}$ seconds) for smaller matrices ($n \le 3$).
+As the matrix size increases, the performance of the parallel execution degrades, and the execution times become increasingly slower.
+This is in contrast to the determinant function, where the parallel execution times increased, but the performance was still somewhat decent for smaller matrices.
+Here, the performance gap between parallel and serial execution widens as the matrix sizes increase, suggesting that the overhead associated with parallelization becomes more pronounced for larger matrices.
+\\\\
+The standard deviation for serial execution times also rises with matrix size.
+This indicates an increased variability in execution times for larger matrices in the serial execution.
+In contrast, the standard deviation for parallel execution times remains relatively stable as matrix size increases, indicating more consistent results across multiple runs.
 \begin{figure}[H]

### 2023-10-17 latex/paper/main.tex
abstract and conclusion

@@
 \section{Abstract}
+Our study explores the performance of distributed linear algebra operations executed in both serial and parallel modes across varying matrix sizes.
+We find that serial execution outperforms parallel execution, especially for smaller matrices, but as matrix size increases, the performance gap widens.
+This suggests that parallelization's overhead can outweigh its advantages for larger matrices.
+Additionally, standard deviation in execution times increases with matrix size.
+Our case study uses the Ray framework on a 10-computer cluster provided by the University of Messina.
+The cluster's 8GB of RAM supports our comprehensive tests of determinant, dot product, matrix inversion, and rank calculations across different matrix sizes.
+This research contributes insights for optimizing distributed computing in the era of big data and scalable computation.
 
@@
 \subsection{Inverse}
+In the parallel execution, the execution times for the inverse function are generally faster for smaller matrices ($n \le 3$), with execution times measured in seconds.
+However, for larger matrices ($n \ge 4$), the parallel execution times increase significantly, with a noticeable increase in execution times and less consistent performance.
+The standard deviation is also considerably higher for parallel execution in larger matrices.
+\\\\
+In contrast, the serial execution times for the inverse function remain relatively stable and are consistently faster than the corresponding parallel execution times for all matrix sizes.
+The standard deviation for serial execution is low.
 
@@
 \subsection{Rank}
+In the case of rank calculation, both parallel and serial show more or less the same amount of execution time, with a linear increase with the matrix size.
+However, the standard deviation for parallel execution increases considerably as the matrix size grows.
+This indicates that parallel execution becomes more variable and less consistent for larger matrices, while serial execution maintains a more stable and predictable performance.
+
 \begin{figure}[H]
@@
 \section{Conclusion}
+Our investigation into the performance of distributed linear algebra operations, executed in both serial and parallel modes, has given important insights into the dynamics of distributed computing across diferent matrix sizes.
+
+First, we observed that serial execution times, when measured in microseconds for smaller matrices ($n \le 3$), consistently outperform their parallel counterparts.
+However, as matrix size increases, the execution times for serial calculations grow.
+This trend agrees with the observation that parallel execution generally lags behind serial execution, and this performance gap widens with growing matrix sizes.
+This suggests that the overhead linked to parallelization can, in fact, eclipse the advantages of parallelism, especially for bigger matrices.
+Thus, we identified that the standard deviation of both serial and parallel execution times increase with matrix size.
+Altough the degree of this increase changes with the operation performed.
+
+The standard deviation for serial execution in the dot product function also exhibits an upward trajectory with growing matrix size.
+In contrast, parallel execution standard deviation remains comparatively constant.
+
+In sum, for our experiemnts, serial execution always outperforms parallel execution, with both serial execution times and standard deviations sensitive to matrix size.
+This reveals that the advantages of parallelization are not universally applicable, with the performance gains being countered by overhead in some scenarios.
 

### 2023-10-17 latex/paper/main.tex
abstract refining

@@
 Our case study uses the Ray framework on a 10-computer cluster provided by the University of Messina.
-The cluster's 8GB of RAM supports our comprehensive tests of determinant, dot product, matrix inversion, and rank calculations across different matrix sizes.
-This research contributes insights for optimizing distributed computing in the era of big data and scalable computation.
+The cluster performs our tests of determinant, dot product, matrix inversion, and rank calculations across different matrix sizes.
 

### 2023-10-17 latex/paper/main.tex
conclusion paragraphing

@@
 Our investigation into the performance of distributed linear algebra operations, executed in both serial and parallel modes, has given important insights into the dynamics of distributed computing across diferent matrix sizes.
-
+\\\\
 First, we observed that serial execution times, when measured in microseconds for smaller matrices ($n \le 3$), consistently outperform their parallel counterparts.
@@
 Altough the degree of this increase changes with the operation performed.
-
+\\\\
 The standard deviation for serial execution in the dot product function also exhibits an upward trajectory with growing matrix size.
 In contrast, parallel execution standard deviation remains comparatively constant.
-
+\\\\
 In sum, for our experiemnts, serial execution always outperforms parallel execution, with both serial execution times and standard deviations sensitive to matrix size.

### 2023-10-30 latex/presentation/main.tex
project reformat
determinant presentation

@@
 \usetheme{metropolis}
-\usepackage{verbatim}
+\usepackage{float}
+\usepackage{listings}
+\usepackage{xcolor}
+\usepackage{transparent}
+
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+% \definecolor{backcolour}{rgb}{0.95,0.95,0.95}
+\definecolor{backcolour}{rgb}{0.98,0.98,0.98}
+
+\lstdefinestyle{mystyle}{
+    backgroundcolor=\color{backcolour},   
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\ttfamily\fontsize{7}{5}\selectfont
+    aboveskip=15pt, % Adjust the space above the listing
+    belowskip=15pt, % Adjust the space below the listing
+    breakatwhitespace=false,         
+    breaklines=true,                 
+    captionpos=b,                    
+    keepspaces=true,                 
+    numbers=left,                    
+    numbersep=5pt,                  
+    showspaces=false,                
+    showstringspaces=false,
+    showtabs=false,                  
+    tabsize=1
+}
+
+\lstset{style=mystyle}
 
@@
 \subtitle{Distributed systems and networks laboratory}
-\author{
-    Gabriele Aloisio [503264] \and
-    Samuel Giacomo Raffa [matricola]
-    }
+\author{Gabriele Aloisio [503264]}
 
@@
 \section{Implementation in Python}
-\begin{frame}{The Matrix and RayMatrix classes}
+\begin{frame}{The Matrix classe}
     Before implementing the parallelized algorithms for Ray, we wrote the code for the serial execution in the \alert{Matrix} class:
@@
 \begin{frame}[fragile]{The Matrix class}
-    \begin{verbatim}
+    \begin{lstlisting}[language=Python]
 class Matrix:
-            def __init__(self, elements):
+    def __init__(self, data):
+        if isinstance(data, list):
+            self.data = [[round(val, 8) for val in row]
+                                        for row in data]
+        else:
+            raise ValueError("Input must be a list")
     [...]
-                self.elements = elements
+    def dot(A, B): [...]
+    def det(self): [...]
+    def rank(self): [...]
     def inv(self): [...]
-            def det(self): [...]
-            def product(self, b): [...]
-            def rank(self, order): [...]
-            [...]
-
-    \end{verbatim}
+\end{lstlisting}
 
@@
 
-\begin{frame}{The RayMatrix class}
-    The \alert{RayMatrix} class is an extension of the Matrix class that inherits its attributes and methods.
-Operations that had to be parallelized were overrided and new auxiliary methods (methods that start with \textit{task\textunderscore}) decorated with \textbf{@ray.remote} were added:
-\end{frame}
-
-\begin{frame}[fragile]{The RayMatrix class}
-    \begin{tiny}
-        \begin{verbatim}
-        class RayMatrix(Matrix):
-            def __init__(self, elements):
-                super().__init__(elements)
-
-            @ray.remote
-            def task_rank_det(submatrix, j):
-
-            @ray.remote
-            def task_get_square_submatrix(self, start_row, start_col, row, order):
-
-            @ray.remote
-            def task_det(self, elements, i):
-
-            @ray.remote
-            def task_inv_cof(self, a, i, j):
-
-            @staticmethod
-            @ray.remote
-            def task_multiply(a, b, i, j, k):
-
-            @staticmethod
-            @ray.remote
-            def task_sum(results):
-
-    \end{verbatim}
-    \end{tiny}
+\begin{frame}{The Matrix class}
+    We modify the Matrix class in a later time to parallelize its operations
 \end{frame}
@@
 
-\begin{frame}[fragile]{Multiplication with Ray (1)}
-    \begin{small}
-        \begin{verbatim}
+\begin{frame}[fragile]{Multiplication with Ray (Serial implementation)}
+    \begin{lstlisting}[language=Python]
 @staticmethod
-def product(a, b):
-    a_columns = a.size()["columns"]
-    a_rows = a.size()["rows"]
+def dot(A, B):
+    if A.is_matrix() and B.is_matrix():
+        a_rows, a_cols = A.shape()
+        b_rows, b_cols = B.shape()
 
-    b_columns = b.size()["columns"]
-
-    if a_rows != b_columns:
+        if a_cols == b_rows:
+            result = [[0] * b_cols for _ in range(A.shape()[0])]
+            for i in range(A.shape()[0]):
+                for j in range(b_cols):
+                    for k in range(a_cols):
+                        result[i][j] += A.data[i][k] * B.data[k][j]
+            return Matrix(result)
+        else:
             raise ValueError(
-            """
-            Number of columns of the first matrix 
-            must match the number of rows of the second matrix
-            """
-            )
+                "Matrix dimensions do not match for dot product")
     else:
-    ...
-
-    \end{verbatim}
-    \end{small}
+        raise ValueError("Dot product requires a Matrix object")
+    \end{lstlisting}
 \end{frame}
 
-\begin{frame}[fragile]{Multiplication with Ray (2)}
-    \begin{tiny}
-        \begin{verbatim}
-a_elements = a.get()  # Get the elements of matrix a
-b_elements = b.get()  # Get the elements of matrix b
+\begin{frame}[fragile]{Multiplication with Ray (dot\_calc)}
+    To parallelize this function, we simply delegate the calculation to an auxiliary function \textit{dot\_calc}:
 
-result = [[0] * b_columns for _ in range(a_rows)]  # Create a result matrix filled with zeros
-tasks = []  # Initialize an empty list to store the tasks
-
-# Iterate over rows of matrix a
-for i in range(a_rows):
-    # Iterate over columns of matrix b
-    for j in range(b_columns):
-        elements_to_multiply = []  # Initialize an empty list to store the elements to be multiplied
-
-        # Iterate over columns of matrix a and rows of matrix b
-        for k in range(a_columns):
-            # Add a task to multiply the elements at position (i, k) and (k, j)
-            elements_to_multiply.append(
-                RayMatrix.task_multiply.remote(a=a_elements, b=b_elements, i=i, j=j, k=k)
-            )
-
-        # Add the task to compute the sum of multiplied elements
-        tasks.append(RayMatrix.task_sum.remote(results=ray.get(elements_to_multiply)))
-    \end{verbatim}
-    \end{tiny}
-\end{frame}
-
-\begin{frame}[fragile]{Multiplication with Ray (3)}
-    \begin{verbatim}
-# Get the results of all the tasks
-results = ray.get(tasks)
-
-# Fill the result matrix with the computed results
-for i in range(a_rows):
-    for j in range(b_columns):
-        result[i][j] = results.pop(0)
-
-return RayMatrix(result)
-
-    \end{verbatim}
-\end{frame}
-
-\begin{frame}[fragile]{Multiplication with Ray(4)}
-    \begin{verbatim}
+    \begin{lstlisting}[language=Python]
 @staticmethod
@@
     return a[i][k] * b[k][j]
-    \end{verbatim}
+    \end{lstlisting}
+
+\end{frame}
+
+\begin{frame}[fragile]{Multiplication with Ray (Parallel)}
+    \begin{lstlisting}[language=Python]
+@staticmethod
+def dot(A, B):
+    if A.is_matrix() and B.is_matrix():
+        a_rows, a_cols = A.shape()
+        b_rows, b_cols = B.shape()
+
+        if a_cols == b_rows:
+            result = [[0] * b_cols for _ in range(a_rows)]
+            futures = []
+
+            for i in range(a_rows):
+                for j in range(b_cols):
+                    for k in range(a_cols):
+                        futures.append(((i, j), t.dot_calc.remote(A, B, i, j, k)))
+
+            for future in futures:
+                i, j = future[0]
+                result[i][j] += ray.get(future[1])
+
+            return Matrix(result)
+        else:
+            raise ValueError(
+                "Matrix dimensions do not match for dot product")
+    else:
+        raise ValueError("Dot product requires a Matrix object")
+\end{lstlisting}
 \end{frame}
@@
 
-\begin{frame}[fragile]{Determinant with Ray (1)}
-    \begin{tiny}
-        \begin{verbatim}
+\begin{frame}[fragile]{Determinant with Ray (Serial)}
+    \begin{lstlisting}[language=Python]
 def det(self):
@@
 
-            \end{verbatim}
-    \end{tiny}
+            \end{lstlisting}
 \end{frame}
 
-\begin{frame}[fragile]{Determinant with Ray (2)}
+\begin{frame}[fragile]{Determinant with Ray (minor)}
+    The \textit{minor} function
+    \begin{lstlisting}[language=Python]
+    def minor(self, i, j):
+    '''
+    Extract a minor matrix by removing the ith row and jth column
+    '''
 
-    \begin{small}
-        \begin{verbatim}
+    data = self.get()
+    minor_data = [row[:j] + row[j + 1:]
+                    for row_idx, row in enumerate(data) if row_idx != i]
+
+    return Matrix(minor_data)
+\end{lstlisting}
+    Then becomes the parallelized function \textit{dist\_minor}:
+\end{frame}
+
+\begin{frame}[fragile]{Determinant with Ray (dist\_minor)}
+    \begin{lstlisting}[language=Python]
 @ray.remote
-def task_det(self, elements, i):
-    size = len(elements)
-    submatrix_det_sum = 0
-    mats = self.get_square_submatrices(i)
+def dist_minor(A, i, j):
+    '''
+    Extract a minor matrix by removing the ith row and jth column
+    '''
 
-    for j in range(1, size):
-        submatrix_det = mats[j].det2()
-        submatrix_det_sum += 
-            elements[i][j] * ((-1) ** (i + j + 2)) 
-                * submatrix_det
+    data = A.get()
+    minor_data = [row[:j] + row[j + 1:]
+                    for row_idx, row in enumerate(data) if row_idx != i]
 
-    return submatrix_det_sum
-    \end{verbatim}
-    \end{small}
+    return Matrix(minor_data)
+        \end{lstlisting}
+    So finally the distributed determinant function will be:
+\end{frame}
 
+\begin{frame}[fragile]{Determinant with Ray (Parallel)}
+
+    \begin{lstlisting}[language=Python]
+    def det(self):
+        if self.is_square():
+            data = self.get()
+            _, cols = self.shape()
+
+            if cols == 1:
+                return data[0][0]
+
+            elif cols == 2:
+                return (data[0][0] * data[1][1]) - (data[0][1] * data[1][0])
+
+            else:
+                det_value = 0
+
+                minors_futures = [Matrix.dist_minor.remote(self, 0, j) for j in range(cols)]
+                minors = ray.get(minors_futures)
+
+                for minor, j in zip(minors, range(cols)):
+                    # minor = self.minor(0, j)
+                    det_value += ((-1) ** j) * data[0][j] * minor.det()
+
+                return det_value
+    \end{lstlisting}
 \end{frame}
@@
 \begin{frame}[fragile]{Inverse with Ray}
-    \begin{small}
-        \begin{verbatim}
+    \begin{lstlisting}[language=Python]
             codice inversa
-        \end{verbatim}
-    \end{small}
+        \end{lstlisting}
 \end{frame}
@@
 \begin{frame}[fragile]{Rank with Ray}
-    \begin{small}
-        \begin{verbatim}
+    \begin{lstlisting}[language=Python]
             codice rango
-        \end{verbatim}
-    \end{small}
+        \end{lstlisting}
 \end{frame}

### 2023-10-30 latex/presentation/main.tex
inverse presentation

@@
 
-\begin{frame}[fragile]{Determinant with Ray (Serial)}
+\begin{frame}[fragile]{Determinant function (Serial)}
     \begin{lstlisting}[language=Python]
@@
 
-\begin{frame}[fragile]{Determinant with Ray (minor)}
+\begin{frame}[fragile]{Determinant function (minor)}
     The \textit{minor} function
@@
 
-\begin{frame}[fragile]{Determinant with Ray (dist\_minor)}
+\begin{frame}[fragile]{Determinant function (dist\_minor)}
     \begin{lstlisting}[language=Python]
@@
 
-\begin{frame}[fragile]{Determinant with Ray (Parallel)}
+\begin{frame}[fragile]{Determinant function (Parallel)}
 
@@
 
-\begin{frame}[fragile]{Inverse with Ray}
+\begin{frame}[fragile]{Inverse function (Serial) (1)}
     \begin{lstlisting}[language=Python]
-            codice inversa
+def inv(self):
+if not self.is_square():
+    raise Exception("Matrix must be square")
+
+elif self.det() == 0:
+    raise Exception("Matrix is not invertible")
+
+else:
+    rows, cols = self.shape()
+    data = self.get()
+
+    det = self.det()
+
+    # special case for 2x2 matrix:
+    if rows == cols == 2:
+        return [[data[1][1] / det, -1 * data[0][1] / det],
+                [-1 * data[1][0] / det, data[0][0] / det]]
+
+...
+        \end{lstlisting}
+\end{frame}
+
+\begin{frame}[fragile]{Inverse function (Serial) (2)}
+    \begin{lstlisting}[language=Python]
+...
+    # find matrix of cofactors
+    cof_matrix = []
+
+    for row in range(rows):
+        cof_row = []
+
+        for column in range(cols):
+            minor = self.minor(row, column)
+
+            cof_row.append(((-1)**(row + column)) * minor.det())
+
+        cof_matrix.append(cof_row)
+
+    cof_matrix = Matrix(cof_matrix).transpose()
+
+    cof_rows, cof_cols = cof_matrix.shape()
+    cof_data = cof_matrix.get()
+
+    for row in range(cof_rows):
+        for column in range(cof_cols):
+            cof_data[row][column] = cof_data[row][column] / det
+
+    return cof_matrix
+        \end{lstlisting}
+\end{frame}
+
+\begin{frame}[fragile]{Distributing the inverse function}
+    To distribute the inverse function we will need two auxiliary functions: \textit{inv\_cof\_matrix} and \textit{inv\_calc}
+\end{frame}
+
+\begin{frame}[fragile]{inv\_cof\_matrix}
+    Find all the cofactor matrices of a given matrix
+    \begin{lstlisting}[language=Python]
+@ray.remote
+def inv_cof_matrix(A, row, cols):
+    from matrix import Matrix
+
+    cof_row = []
+    minor_futures = [Matrix.dist_minor.remote(A, row, col) for col in range(cols)]
+    minors = ray.get(minor_futures)
+
+    for col,minor in zip(range(cols), minors):
+        cof_row.append(((-1)**(row + col)) * minor.det())
+
+    return cof_row
+        \end{lstlisting}
+\end{frame}
+
+\begin{frame}[fragile]{inv\_calc}
+    Calculation step for the main function
+    \begin{lstlisting}[language=Python]
+@ray.remote
+def inv_calc(row, cof_cols, cof_data, det):
+    for col in range(cof_cols):
+        return (row, col, (cof_data[row][col] / det))
+        \end{lstlisting}
+\end{frame}
+
+\begin{frame}[fragile]{Inverse function (Parallel) (1)}
+    \begin{lstlisting}[language=Python]
+def inv(self):
+if not self.is_square():
+    raise Exception("Matrix must be square")
+
+elif self.det() == 0:
+    raise Exception("Matrix is not invertible")
+
+else:
+    rows, cols = self.shape()
+    data = self.get()
+    det = self.det()
+
+    # special case for 2x2 matrix:
+    if rows == cols == 2:
+        m = [[data[1][1] / det, -1 * data[0][1] / det],
+                [-1 * data[1][0] / det, data[0][0] / det]]
+
+        return Matrix([[round(i, 8) for i in j] for j in m])
+...    
+            \end{lstlisting}
+\end{frame}
+
+\begin{frame}[fragile]{Inverse function (Parallel) (2)}
+    \begin{lstlisting}[language=Python]
+...
+    # find matrix of cofactors
+    cof_rows_futures = [t.inv_cof_matrix.remote(self, row, cols) for row in range(rows)]
+    cof_matrix = ray.get(cof_rows_futures)
+    cof_matrix = Matrix(cof_matrix).transpose()
+
+    cof_rows, cof_cols = cof_matrix.shape()
+    cof_data = cof_matrix.get()
+
+    data_futures = [t.inv_calc.remote(row, cof_cols, cof_data, det) for row in range(cof_rows)]
+
+    for data in ray.get(data_futures):
+        for row, col, value in [data]:
+            cof_data[row][col] = value
+
+    return Matrix([[round(i, 8) for i in j] for j in cof_matrix])
             \end{lstlisting}
@@
 
-\begin{frame}[fragile]{Rank with Ray}
+\begin{frame}[fragile]{Rank function}
     \begin{lstlisting}[language=Python]

### 2023-10-30 latex/presentation/main.tex
rank presentation

@@
 
-\begin{frame}[fragile]{Rank function}
+\begin{frame}[fragile]{Rank function (Serial)}
     \begin{lstlisting}[language=Python]
-            codice rango
+def rank(self):
+rows, cols = self.shape()
+j1 = min(rows, cols)
+
+for i in range(j1, 1, -1):
+    if self.get_square_submatrices(i) != -1:
+        return i
+        \end{lstlisting}
+\end{frame}
+
+\begin{frame}[fragile]{get\_square\_submatrices (Serial)}
+    The main protagonist here is the \textit{get\_square\_submatrices} function:
+    \begin{lstlisting}[language=Python]
+def get_square_submatrices(self, order):
+'''
+Auxiliary function for the rank() function
+'''
+data = self.get()
+rows, cols = self.shape()
+submatrices = []
+
+for start_row in range(rows - order + 1):
+    for start_col in range(cols - order + 1):
+        submatrix = []
+
+        for row in range(order):
+            submatrix.append(
+                data[start_row + row][start_col:start_col + order])
+
+        submatrices.append(Matrix(submatrix))
+
+return submatrices
+                    \end{lstlisting}
+\end{frame}
+
+\begin{frame}[fragile]{{get\_submatrix\_task}}
+    We will distribute \textit{get\_square\_submatrices} by defining an auxiliary function \textit{get\_submatrix\_task}:
+
+    \begin{lstlisting}[language=Python]
+@ray.remote
+def get_submatrix_task(start_row, start_col, order, data):
+        from matrix import Matrix
+        submatrix = []
+
+        for row in range(order):
+            submatrix.append(
+                data[start_row + row][start_col:start_col + order])
+
+        return submatrix
+        \end{lstlisting}
+\end{frame}
+\begin{frame}[fragile]{get\_square\_submatrices (Parallel)}
+    \begin{lstlisting}[language=Python]
+def get_square_submatrices(self, order):
+'''
+Auxiliary function for the rank() function
+'''
+data = self.get()
+rows, cols = self.shape()
+futures = []
+
+for start_row in range(rows - order + 1):
+    for start_col in range(cols - order + 1):            
+        futures.append(t.get_submatrix_task.remote(start_row, start_col, order, data))
+
+return ray.get(futures) 
         \end{lstlisting}

### 2023-11-08 latex/paper/main.tex
finished slides

@@
 \begin{figure}[H]
-    \includegraphics[width=0.8\textwidth, keepaspectratio]{img/det_2_10_3.png}
+    \includegraphics[width=0.8\textwidth, keepaspectratio]{../img/det_2_10_3.png}
     \centering
@@
 
-    \includegraphics[width=0.8\textwidth, keepaspectratio]{img/det_2_10_3_detail.png}
+    \includegraphics[width=0.8\textwidth, keepaspectratio]{../img/det_2_10_3_detail.png}
     \centering
@@
 \begin{figure}[H]
-    \includegraphics[width=0.8\textwidth, keepaspectratio]{img/dot_2_30_5.png}
+    \includegraphics[width=0.8\textwidth, keepaspectratio]{../img/dot_2_30_5.png}
     \centering
@@
 
-    \includegraphics[width=0.8\textwidth, keepaspectratio]{img/dot_2_30_5_detail.png}
+    \includegraphics[width=0.8\textwidth, keepaspectratio]{../img/dot_2_30_5_detail.png}
     \centering
@@
 \begin{figure}[H]
-    \includegraphics[width=0.8\textwidth, keepaspectratio]{img/inv_2_8_3.png}
+    \includegraphics[width=0.8\textwidth, keepaspectratio]{../img/inv_2_8_3.png}
     \centering
@@
 
-    \includegraphics[width=0.8\textwidth, keepaspectratio]{img/inv_2_8_3_detail.png}
+    \includegraphics[width=0.8\textwidth, keepaspectratio]{../img/inv_2_8_3_detail.png}
     \centering
@@
 \begin{figure}[H]
-    \includegraphics[width=0.8\textwidth, keepaspectratio]{img/rank_2_500_5.png}
+    \includegraphics[width=0.8\textwidth, keepaspectratio]{../img/rank_2_500_5.png}
     \centering

### 2023-11-08 latex/presentation/main.tex
finished slides

@@
 \usepackage{transparent}
+\usepackage{wasysym}
 
@@
 \end{frame}
+\section{Results of tests}
+\begin{frame}[fragile]{Description of tests}
+    For our experiments, we considered the following parameters:
+\begin{itemize}
+    \item \textbf{Determinant calculation:} $n = 2 \dots 10$, each $n$ executed three times.
+    \item \textbf{Dot product:} $n = 2 \dots 30$, each $n$ executed five times.
+    \item \textbf{Inverse:} $n = 2 \dots 8$, each $n$ executed three times.
+    \item \textbf{Rank:} $n = 2 \dots 500$, each $n$ executed five times.
+\end{itemize}
+\vfill
+For each type of operation, the code iterates through different matrix sizes, executing the operation multiple times for each size to capture variations in performance
+\end{frame}
+
+\begin{frame}[fragile]{Results of tests (Multiplication)}
+    \centering
+    \includegraphics[width=\textwidth]{../img/dot_2_30_5.png}
+\end{frame}
+
+\begin{frame}[fragile]{Results of tests (Determinant)}
+    \centering
+    \includegraphics[width=\textwidth]{../img/det_2_10_3.png}
+\end{frame}
+
+\begin{frame}[fragile]{Results of tests (Inverse)}
+    \centering
+    \includegraphics[width=\textwidth]{../img/inv_2_8_3.png}
+\end{frame}
+
+\begin{frame}[fragile]{Results of tests (Rank)}
+    \centering
+    \includegraphics[width=\textwidth]{../img/rank_2_500_5.png}
+\end{frame}
+
+\section{Conclusion}
+\begin{frame}[fragile]{Conclusion}
+    Serial execution always outperforms parallel execution, with both serial execution times and standard deviations sensitive to matrix size.
+This reveals that the advantages of parallelization are not universally applicable, with the performance gains being countered by overhead in some scenarios.
+\end{frame}
+
+\begin{frame}[standout]
+    Thanks for reading
+    \\
+    \Huge\smiley{}
+\end{frame}
 
